\section{Zariski Dense Subsets}


\begin{fluff}
  For this section let $k$ be an infinite field.
\end{fluff}


\begin{definition}
  \label{definition: Zariski density}
  Let $V$ be a finite-dimensional $k$-vector space.
  A subset $X \subseteq V$ is \emph{Zariski dense \textup(over $k$\textup)} if for every polynomial map $f \colon V \to k$ we have that
  \[
              \restrict{f}{X} = 0
    \implies  f = 0 \,.
  \]
  If $X \subseteq Y \subseteq V$ then \emph{$X$ is Zariski dense in $Y$ \textup(over $k$\textup)} if for every polynomial function $f \colon V \to k$ we have that
  \[
              \restrict{f}{X} = 0
    \implies  \restrict{f}{Y} = 0 \,.
  \]
\end{definition}


\begin{example}
  Let $V$ be a finite-dimensional $k$-vector space.
  \begin{enumerate}
    \item 
      Any infinite subset $X \subseteq k$ is Zariski dense:
      Let $f \colon V \to k$ be a polynomial function with $\restrict{f}{X} = 0$.
      There exists some polynomial $p \in k[X]$ with $f(\lambda) = p(\lambda)$ for all $\lambda \in k$ because $f$ is polynomial.
      It follows from $\restrict{f}{X} = 0$ that every $x \in X$ is a zero of $p$, which shows that $p$ has infinitley many zeroes.
      This can only be the case for $p = 0$, and thus $f = 0$.
    \item
      Let $U \subsetneq V$ be a proper $k$-linear subspace.
      Then $U$ is not Zariski dense in $V$ over $k$:
      To see this let $v_1, \dotsc, v_m, v_{m+1}, \dotsc, v_n$ be a $k$-basis of $V$ such that $v_1, \dotsc, v_m$ is a $k$-basis of $U$.
      Then $m < n$ because $U$ is a proper subspace of $V$.
      Let $\pi \colon V \to k$ be the projection onto the last coordinate, i.e.
      \[
          \pi\left( \lambda_1 v_1 + \dotsb + \lambda_n v_n \right)
        = \lambda_n
      \]
      for all $\lambda_1, \dotsc, \lambda_n \in k$.
      We then have that $\restrict{\pi}{U} = 0$ but $\pi \neq 0$, which shows that $U$ is not Zariski dense in $V$ over $k$.
    \item
      Let $k$ be a finite field.
      We have seen in Remark~\ref{remark: polynomial functions over infinite fields} that every function $f \colon V \to k$ is a polynomial function.
      It follows that the only Zariski-dense subset $X \subseteq V$ is $X = V$ itself.
  \end{enumerate}
\end{example}


\begin{warning}
  The previous examples show that the notion of Zariski density depend on the choice of the underlying field $k$:
  It follows from the first example that $\Real \subseteq \Complex$ is Zariski dense over $\Complex$, while it follows from the second example that $\Real \subseteq \Complex$ is not Zariski dense over $\Real$.
\end{warning}


\begin{lemma}
  Let $V$ be a finite-dimensional $k$-vector space and let $h \colon V \to k$ be a non-zero polynomial function.
  Then the non-vanishing set
  \[
              V_h
    \defined  \{
                v \in V
              \suchthat
                h(v) \neq 0
              \}
  \]
  is Zariski dense in $V$.
\end{lemma}
\begin{proof}
  Let $f \colon V \to k$ be a polynomial function with $\restrict{f}{V_h} = 0$.
  Then
  \[
      (fh)(v)
    = f(v)h(v)
    = 0
  \]
  for all $v \in V$, and thus $fh = 0$.
  Note that $\mc{P}(V) \cong k[X_1, \dotsc, X_{(\dim V)}]$ is an integral domain.
  It therefore follows from $fh = 0$ and $h \neq 0$ that $f = 0$.
\end{proof}


\begin{corollary}
  \label{corollary: GLn is Zariski dense in Mn}
  The group $\GL_n(k)$ is Zariski dense in $\Mat_n(k)$.
\end{corollary}
\begin{proof}
  The group $\GL_n(k)$ is the non-vanishing set of $\det \colon \Mat_n(k) \to k$.
\end{proof}


\begin{lemma}\label{lemma: zariski density orbits}
  Let $V$ be a finite-dimensional representation of a group $G$ and let $f \colon V \to k$ be a $G$-invariant polynomial function.
  \begin{enumerate}
    \item
      Suppose that $X \subseteq V$ is a subset such that that the orbit
      \[
          G.X
        = \{
            g.x
          \mid
            g \in G,
            x \in X
          \}
      \]
      is Zariski-dense in $V$.
      If $\restrict{f}{X} = 0$ then $ f= 0$.
    \item
      If there exists a Zariski dense $G$-orbit for $V$ then $f$ is constant.
  \end{enumerate}
\end{lemma}
\begin{proof}
  \leavevmode
  \begin{enumerate}
    \item
      It follows from the $G$-invariance of $f$ that $\restrict{f}{G.X} = 0$ because
      \[
          f(g.x)
        = \left( g^{-1}.f \right)(x)
        = f(x)
        = 0
      \]
      for all $g \in G$, $x \in X$.
      It follows that $f = 0$ because $G.X$ is Zariski dense in $V$.
    \item
      Suppose that the orbit of $x \in V$ is Zariski dense in $V$.
      We can then apply the previous part of this lemma to $X = \{x\}$ and the $G$-invariant polynomial function $f - f(x) \colon V \to k$, which shows that $f - f(x) = 0$ and thus $f = f(x)$.
    \qedhere
  \end{enumerate}
\end{proof}


\begin{proposition}
  \label{proposition: diagonalizable matrices are dense for algebraically closed}
  If $k$ is algebraically closed then the set of diagonalizable matrices,
  \[
              \Diag_n(k)
    \defined  \{
                A \in \Mat_n(k)
              \mid
                \text{$A$ is diagonalizable}
              \} \,,
  \]
  is Zariski-dense in $\Mat_n(k)$.
\end{proposition}
\begin{proof}
  Let $f \colon \Mat_n(k) \to k$ with $\restrict{f}{\Diag_n(k)} = 0$ and let $A \in \Mat_n(k)$.
  The matrix $A$ is triangularizable over $k$ because $k$ is algebraically closed, so there exists some $S \in \GL_n(k)$ such that $S A S^{-1}$ is an upper triangular matrix with diagonal entries $b_1, \dotsc, b_n \in k$ (not necessarily pairwise distinct).
  
  We want to \enquote{deform} the matrix $A$ to make it diagonalizable:
  
  Let $a_1, \dotsc, a_n \in k$ be pairwise different (such $a_i$ exist because $k$ is infinite), and consider the map $M \colon k \to \Mat_n(k)$ given by
  \[
              M(z)
    \defined  S^{-1}
              \left(
              S A S^{-1}
              + z
              \begin{pmatrix}
                a_1 &         &     \\
                    & \ddots  &     \\
                    &         & a_n
              \end{pmatrix}
              \right)
              S
  \]
  for all $z \in k$.
  Then $M(0) = A$, so one may think of $M(z)$ as a \enquote{deformation} of $A$ along a parameter $z \in k$.
  Note that $M(z)$ has the eigenvalues $b_1 + a_1 z, \dotsc, b_n + a_n z$ because
  \begin{align*}
      S M(z) S^{-1}
    =     S A S^{-1}
        + z
        \begin{pmatrix}
          a_1 &         &     \\
              & \ddots  &     \\
              &         & a_n
        \end{pmatrix}
    &=  \begin{pmatrix}
          b_1 & \cdots  & *       \\
              & \ddots  & \vdots  \\
              &         & b_n
        \end{pmatrix}
        + z
        \begin{pmatrix}
          a_1 &         &     \\
              & \ddots  &     \\
              &         & a_n
        \end{pmatrix} \\
    &=  \begin{pmatrix}
          b_1 + z a_1 & \cdots  & *           \\
                      & \ddots  & \vdots      \\
                      &         & b_n + z a_n
        \end{pmatrix}.
  \end{align*}
  Any two eigenvalues $b_i + a_i z$ and $b_j + a_j z$ of $M(z)$ coincide for only one value of $z$, namely for $z = (b_j - b_i)/(a_i - a_j)$.
  It follows that $M(z)$ has pairwise different eigenvalues, and is therefore digonalizable, for all but finitely many $z \in k$.
  
  It follows from $\restrict{f}{\Diag_n(k)} = 0$ that $(f \circ M)(z) = 0$ for all but finitely many $z \in k$, and therefore that $f \circ M = 0$.
  For $z = 0$ we find that
  \[
      0
    = (f \circ M)(0)
    = f(M(0))
    = f(A) \,.
  \]
  This shows that $f(A) = 0$ for every $A \in \Mat_n(k)$, i.e.\ that $f = 0$.
\end{proof}


\begin{remark}
  The above proof actually shows that the set of matrices with pairwise different eigenvalues is Zariski dense if $k$ is algebraically closed.
\end{remark}


\begin{corollary}
  \label{corollary: diagonal matrices dense alg closed}
  Let $k$ be an algebraically closed field and let $\GL_n(k)$ act on $\Mat_n(k)$ by conjugation.
  Let $f \in \Mat_n(k) \to k$ be a $\GL_n(k)$-invariant polynomial function and let $D \subseteq \Mat_n(k)$ be the subset of diagonal matrices.
  If $\restrict{f}{D} = 0$ then $f = 0$.
\end{corollary}
\begin{proof}
  This follows from Lemma~\ref{lemma: zariski density orbits} because the orbit $\GL_n(k).D = \Diag_n(k)$ is Zariski dense in $\Mat_n(k)$ by Proposition~\ref{proposition: diagonalizable matrices are dense for algebraically closed}.
\end{proof}





\subsection*{Extension of scalars}


\begin{fluff}
  The first of the Zariski density properties stated in Proposition~\ref{proposition: zariski density properties} has been proven by Corollary~\ref{corollary: GLn is Zariski dense in Mn}, and the second Zariski density property has been proven by Corollary~\ref{corollary: diagonal matrices dense alg closed} for algebraically closed fields.
  
  To show that the second Zariski density property holds for arbitrary infinite fields we now examine how Zariski density changes under extension of scalars.
  
  In the following let $L/k$ be a field extension.
  For every $k$-vector space $V$ we denote by $V_L = L \otimes_k V$ the extension of scalars of $V$, and regard $V$ as a $k$-linear subspace of $V_L$ (see Appendix~\ref{appendix: extension of scalars} for a short introduction to extension of scalars).
  If $V$ is finite dimensional then we abbreviate $\mc{P}(V) = \mc{P}_k(V)$ and $\mc{P}(V_L) = \mc{P}_L(V_L)$.
\end{fluff}


\begin{lemma}
  \label{lemma: product of Zariski dense is again Zariski dense}
  Let $V$, $W$ be $k$-vector spaces and let $X \subseteq V$ and $Y \subseteq W$ be Zariski dense.
  Then $X \times Y \subseteq V \times W$ is again Zariski dense.
\end{lemma}


\begin{proof}
  Let $f \colon V \times W \to k$ be a polynomial function with $\restrict{f}{X \times Y} = 0$.
  We then have that
  \[
      f(x,y)
    = 0
    \qquad
    \text{for all $x \in X$, $y \in Y$} \,.
  \]
  For fixed $x \in X$ the map $f(x,-) \colon W \to k$ is polynomial with $\restrict{f(x,-)}{Y} = 0$, so it follows that $f(x,-) = 0$ because $Y$ is Zariski dense in $W$.
  It follows that
  \[
      f(x,w)
    = 0
    \qquad
    \text{for all $x \in X$, $w \in W$}.
  \]
  For fixed $w \in W$ the map $f(-,w) \colon V \to k$ is polynomial with $\restrict{f(-,w)}{X} = 0$, so it follows that $f(-,w) = 0$ because $X$ is Zariski dense in $V$.
  It follows that
  \[
      f(v,w)
    = 0
    \qquad
    \text{for all $v \in V$, $w \in W$} \,.
  \]
  This shows that $f = 0$.
\end{proof}


\begin{corollary}
\label{corollary: kn is Zariski dense in Ln}
  The subset $k^n \subseteq L^n$ is Zariski dense over $L$.
\end{corollary}


\begin{proof}
  This follows from Lemma~\ref{lemma: product of Zariski dense is again Zariski dense} because $k \subseteq L$ is Zariski dense over $L$ (because $k$ is infinite).
\end{proof}



% \begin{proof}
%   We show the statement by induction over $n \geq 0$.
%   For $n = 0$ the statement holds because $k^n = 0 = L^n$.
%   So let $n \geq 1$, and suppose that the statemen holds for $n-1$.
%   
%   Let $f \in \mc{P}(L^n)$ with $\restrict{f}{k^n} = 0$.
%   There exists some polynomial $p \in L[X_1, \dotsc, X_n]$ with $f((\lambda_1, \dotsc, \lambda_n)) = p(\lambda_1, \dotsc, \lambda_n)$ for all $(\lambda_1, \dotsc, \lambda_n) \in L^n$ because $f$ is a polynomial function.
%   We may write $p$ as
%   \[
%       p(X_1, \dotsc, X_n)
%     = \sum_{i=0}^\infty p_i(X_1, \dotsc, X_{n-1}) X_n^i
%   \]
%   with $p_i \in L[X_1, \dotsc, X_{n-1}]$ for all $i \geq 0$ and $p_i \neq 0$ for only finitely many $i$.
%   
%   Let $(\lambda_1, \dotsc, \lambda_{n-1}) \in k^{n-1}$ and consider the polynomial $\bar{p} \in L[X_n]$ given by
%   \[
%               \bar{p}(X_n)
%     \coloneqq p(\lambda_1, \dotsc, \lambda_{n-1}, X_n)
%     =         \sum_{i=0}^\infty p_i(\lambda_1, \dotsc, \lambda_{n-1}) X_n^i \,.
%   \]
%   Then $\bar{p}(\lambda_n) = 0$ for all $\lambda_n \in k$, and it follows that $\bar{p} = 0$ (because $\bar{p}$ has infinitely many zeroes).
%   It follows that all coefficients of $\bar{p}$ vanish, i.e.\ that $p_i(\lambda_1, \dotsc, \lambda_{n-1}) = 0$ for all $i \geq 0$, because $L$ is infinite.
%   
%   This shows that the polynomials $p_i$ vanish on all of $k^{n-1}$, which by induction hypothesis implies that they vanish on all of $L^{n-1}$.
%   It follows that $p_i = 0$ for all $i \geq 0$ because $L$ is infinite.
%   This shows that $p = 0$ and thus $f = 0$.
% \end{proof}


\begin{corollary}
  \label{corollary: V is Zariski dense in VL}
  Let $V$ be a finite-dimensional $k$-vector space.
  Then $V$ is Zariski dense in $V_L$ over $L$.
\end{corollary}


\begin{proof}
  Let $v_1, \dotsc, v_n$ be a $k$-basis of $V$.
  Then $1 \otimes v_1, \dotsc, 1 \otimes v_n$ is an $L$-basis of $V_L$.
  The isomorphism of $L$-vector spaces $\varphi \colon L^n \to V_L$ with $\varphi(e_i) = 1 \otimes v_i$ maps $k^n \subseteq L^n$ onto $\varphi(k^n) = V$.
  The statement now follows from Corollary~\ref{corollary: kn is Zariski dense in Ln} because $\varphi$ is a polynomial isomorphism, since both $\varphi$ and $\varphi^{-1}$ are linear and therefore polynomial.
\end{proof}


\begin{proposition}
  \label{proposition: consistent choice of identification for polynomial functions}
  Let $V$ be a finite-dimensional $k$-vector space.
  \begin{enumerate}
    \item
      Every $k$-polynomial map $f \colon V \to k$ extends uniquely to an $L$-polynomial map $\overline{f} \colon V_L \to L$ with
      \[
          \overline{f}(1 \otimes v)
        = f(v)
      \]
      for all $v \in V$.
    \item
      The map $i \colon \mc{P}(V) \to \mc{P}(V_L)$, $f \mapsto \overline{f}$ is a homomorphism of $k$-algebras, which extends to an isomorphism of $L$-algebras
      \[
                                I
        \colon                  \mc{P}(V)_L
        \xrightarrow{\,\sim\,}  \mc{P}(V_L)
      \]
    \item
      Let $v_1, \dotsc, v_n$ be a $k$-basis of $V$ and let $\varphi_1, \dotsc, \varphi_n \in \mc{P}(V)$ be the corresponding coordinate functions.
      Let $\psi_1, \dotsc, \psi_n \in \mc{P}(V_L)$ be the coordinate functions corresponding to the basis $1 \otimes v_1, \dotsc, 1 \otimes v_n$ of $V_L$.
      
      Let $\Phi \colon \mc{P}(V) \to k[X_1, \dotsc, X_n]$ be the unique isomorphism of $k$-algebras with $\Phi(\varphi_i) = X_i$ for all $i$, and let $\Psi \colon \mc{P}(V_L) \to L[X_1, \dotsc, X_n]$ be the unique isomorphism of $L$-algebras with $\Psi(\psi_i) = X_i$ for all $i$.
      
      Then the following diagram commutes:
      \begin{equation}
        \label{equaton: commuting prim}
        \begin{tikzcd}[row sep = large]
            {}
          & \mc{P}(V)
            \arrow{dl}[above left]{\can}
            \arrow{dr}[above right]{i}
            \arrow{dd}[near start, left]{\Phi}
          & {}
          \\
            \mc{P}(V)_L
            \arrow[crossing over]{rr}[near start, above]{I}
            \arrow{dd}[left]{\Phi_L}
          & {}
          & \mc{P}(V_L)
            \arrow{dd}[right]{\Psi}
          \\
            {}
          & k[X_1, \dotsc, X_n]
            \arrow{dl}[above left]{\can}
            \arrow[hook']{dr}
          & {}
          \\
            k[X_1, \dotsc, X_n]_L
            \arrow{rr}[above]{\sim}
          & {}
          & L[X_1, \dotsc, X_n]
        \end{tikzcd}
      \end{equation}
  \end{enumerate}
\end{proposition}


\begin{proof}
    Let $v_1, \dotsc, v_n$ be a $k$-basis of $V$.
    
    There exists a polynomial $p \in k[X_1, \dotsc, X_n]$ with
    \[
        f(\lambda_n v_1 + \dotsb + \lambda_n v_n)
      = p(\lambda_1, \dotsc, \lambda_n)
    \]
    for all $\lambda_1, \dotsc, \lambda_n \in k$ because the the map $f$ is $k$-polynomial.
    We can regard $p$ as a polynomial $p \in L[X_1, \dotsc, X_n]$.
    With respect to the $L$-basis $1 \otimes v_1, \dotsc, 1 \otimes v_n$ of $V_L$ the polynomial $p$ then defines an $L$-polynomial map $\overline{f} \colon V_L \to L$ given by
    \[
        \overline{f}(\lambda_1 (1 \otimes v_1) + \dotsb + \lambda_n (1 \otimes v_n))
      = p(\lambda_1, \dotsc, \lambda_n)
    \]
    for all $\lambda_1, \dotsc, \lambda_n \in L$.
    We have for every $v \in V$ with $v = \sum_{i=1}^n \lambda_i v_i$ that
    \[
        \overline{f}(1 \otimes v)
      = \overline{f}(\lambda_1 (1 \otimes v_1) + \dotsb + \lambda_n (1 \otimes v_n))
      = p(\lambda_1, \dotsc, \lambda_n)
      = f(v) \,.
    \]
    The uniqueness of $\overline{f}$ follows from the Zariski density of $V \subseteq V_L$ over $L$.
    
    Note that the diagram
    \[
      \begin{tikzcd}[sep = large]
          \mc{P}(V)
          \arrow{r}[above]{i}
          \arrow{d}[left]{\Phi}
        & \mc{P}(V_L)
          \arrow{d}[right]{\Psi}
        \\
          k[X_1, \dotsc, X_n]
          \arrow[hook]{r}
        & L[X_1, \dotsc, X_n]
      \end{tikzcd}
    \]
    commutes by the above construction of $i$.
    It follows that
    \[
                                  i
      \colon                      \mc{P}(V)
      \xrightarrow{\,\Phi\,}      k[X_1, \dotsc, X_n]
      \hookrightarrow             L[X_1, \dotsc, X_n]
      \xrightarrow{\,\Psi^{-1}\,} \mc{P}(V_L)
    \]
    is a composition of $k$-algebra homomorphisms, and thus a $k$-algebra homomorphism itself.
    It follows (from Lemma~\ref{lemma: universal property of extension of scalars for algebras}) that $i$ extends uniquely to an $L$-algebra homomorphism $I \colon \mc{P}(V)_L \to \mc{P}(V_L)$ such that the diagram
    \[
      \begin{tikzcd}[sep = large]
          \mc{P}(V)_L
          \arrow[dashed]{r}[above]{I}
        & \mc{P}(V_L)
        \\
          \mc{P}(V)
          \arrow{u}[left]{\can}
          \arrow{ru}[below right]{i}
        & {}
      \end{tikzcd}
    \]
    commutes.
    
    For the commutativity of the diagram~\eqref{equaton: commuting prim} we note that on elements we get the following diagram, which does commute:
    \[
      \begin{tikzcd}[column sep = large]
          {}
        & \varphi_i
          \arrow[mapsto]{dl}
          \arrow[mapsto]{dr}
          \arrow[mapsto]{dd}
        & {}
        \\
          1 \otimes \varphi_i
          \arrow[crossing over, mapsto]{rr}
          \arrow[mapsto]{dd}
        & {}
        & \psi_i
          \arrow[mapsto]{dd}
        \\
          {}
        & X_i
          \arrow[mapsto]{dl}
          \arrow[mapsto]{dr}
        & {}
        \\
          1 \otimes X_i
          \arrow[mapsto]{rr}
        & {}
        & X_i
      \end{tikzcd}
    \]
    It follows that the diagram~\eqref{equaton: commuting prim} commutes because the occuring maps are all algebra homomorphisms and these elements generate their respective algebras (as $L$-algebras for the upper four in the front, and as $k$-algebras for the two in the back).
    
    It remains to show that $I$ is an isomorphism.
    This follows from the commutativity of the diagram~\eqref{equaton: commuting prim} because
    \[
                                  I
      \colon                      \mc{P}(V)_L
      \xrightarrow{\,\Phi_L\,}    k[X_1, \dotsc, X_n]_L
      \xrightarrow{\,\sim\,}      L[X_1, \dotsc, X_n]
      \xrightarrow{\,\Phi^{-1}\,} \mc{P}(V_L)
    \]
    is a composition of isomorphisms.
\end{proof}


\begin{definition}
  Let $V$ be a finite-dimensional $k$-vector space.
  \begin{enumerate}
    \item
      For $X \subseteq V$ the \emph{vanishing ideal of $X$} is given by
      \[
        \mc{I}_k(X)
        \defined  \{
                    f \in \mc{P}_k(V)
                  \suchthat
                    \text{$f(x) = 0$ for all $x \in X$}
                  \} \,.
      \]
      We also write $\mc{I}(X)$ instead of $\mc{I}_k(X)$ if the field $k$ is clear from the context.
    \item
      For every point $a \in V$ we set
      \[
                  \mf{m}_a
        \defined  \mc{I}(\{a\})
        =         \{
                    f \in \mc{P}(V)
                  \suchthat
                    f(a) = 0
                  \} \,.
      \]
  \end{enumerate}
\end{definition}


\begin{lemma}
  Let $V$ be a finite-dimensional $k$-vector space.
  \begin{enumerate}
    \item
      For $X \subseteq V$ the subset $\mc{I}(X) \subseteq \mc{P}(V)$ is an ideal.
    \item
      If $X \subseteq Y \subseteq V$ then $\mc{I}(Y) \subseteq \mc{I}(X)$.
      Furthermore, $X$ is Zariski-dense in $Y$ if and only if $\mc{I}(X) = \mc{I}(Y)$.
    \item
      Let $\{X_i\}_{i \in I}$ be a collection of subsets $X_i \subseteq V$. Then
      \[
          \mc{I}\left( \bigcup_{i \in I} X_i \right)
        = \bigcap_{i \in I} \mc{I}(X_i) \,.
      \]
  \end{enumerate}
\end{lemma}


\begin{lemma}
  \label{lemma: maximal ideal correspondin to a point}
  For $a = (a_1, \dotsc, a_n) \in k^n$ the ideal $\mf{m}_a$ is maximal and given by
  \[
    \mf{m}_a = (X_1 - a_1, \dotsc, X_n - a_n) \,,
  \]
  where we identify $\mc{P}(k^n)$ with $k[X_1, \dotsc, X_n]$ with  as explained in~\ref{fluff: identify polynomials with polynomial maps for kn}.
\end{lemma}


\begin{proof}
  The evaluation map
  \[
            \varepsilon_a
    \colon  k[X_1, \dotsc, X_n]
    \to     k,
    \quad   f
    \mapsto f(a)
  \]
  is a surjective homomorphism of $k$-algebras with $\ker \varepsilon_a = \mf{m}_a$.
  It follows that $\mf{m}_a$ is a maximal ideal because $k[X_1, \dotsc, X_n]/\mf{m}_a \cong k$ is a field.
  
  The ideal $\mf{m} \defined (X_1 - a_1, \dotsc, X_n - a_n)$ is also maximal:
  
  We consider first the case $a_1 = \dotsb = a_n = 0$.
  Then $\mf{m} = (X_1, \dotsc, X_n)$ has a $k$-basis given by all monomials $X_1^{\alpha_1} \dotsm X_n^{\alpha_n} \neq 1$.
  It follows that the $k$-algebra $k[X_1, \dotsc, X_n]/\mf{m}$ has a basis given by the single element $\overline{1}$, and is therefore one-dimensional.
  It follows that $k[X_1, \dotsc, X_n]/\mf{m} \cong k$ as $k$-algebras.
  Then $\mf{m}$ is maximal because $k$ is a field.
  
  For general $a \in k^n$ we observe that there exists an automorphism of $k$-algebras $\Phi \colon k[X_1, \dotsc, X_n] \to k[X_1, \dotsc, X_n]$ with $\Phi(X_i) =  X_i + a_i$ for all $i = 1, \dotsc, n$.
  Then $\Phi(\mf{m}) = (X_1, \dotsc, X_n)$ is maximal by the above argumentation, so $\mf{m}$ itself is also maximal.
  
  It follows that $\mf{m}_a = \mf{m}$ because both ideals are maximal and $\mf{m} \subseteq \mf{m}_a$.
\end{proof}


\begin{corollary}
  Let $V$ be a finite-dimensional $k$-vector space and $X \subseteq V \subseteq V_L$.
  Then
  \[
      \mc{I}_k(X)_L
    = \mc{I}_L(X) \,,
  \]
  where we identify $\mc{P}(V)_L$ with $\mc{P}(V_L)$ as explained in Proposition~\ref{proposition: consistent choice of identification for polynomial functions}.
\end{corollary}
\begin{proof}
  By choosing a basis of $V$ we may identify in a consistent way
  \begin{itemize}
    \item
      the $k$-vector space $V$ with $k^n$,
    \item
      the $L$-vector space $V_L$ with $L^n$,
    \item
      the $k$-algebra $\mc{P}(V)$ with $k[X_1, \dotsc, X_n]$,
    \item
      the $L$-algebras $\mc{P}(V)_L$ and $\mc{P}(V_L)$ with $L[X_1, \dotsc, X_n]$
  \end{itemize}
  by Proposition~\ref{proposition: consistent choice of identification for polynomial functions}.
  For every point $a \in k^n$ we then have that
  \begin{align*}
        \mc{I}_k(\{a\})_L
    &=  L \otimes_k \mc{I}_k(\{a\}) \\
    &=  L \otimes_k ( X_1 - a_1, \dotsc, X_n - a_n )_{ k[X_1, \dotsc, X_n] } \\
    &=  ( X_1 - a_1, \dotsc, X_n - a_n )_{ L[X_1, \dotsc, X_n] }  \\
    &=  \mc{I}_L(\{a\}) \,.
  \end{align*}
  For every subset $X \subseteq W$ we therefore have that
  \begin{align*}
        \mc{I}_k(X)_L
    &=  L \otimes_k \mc{I}_k(X)
     =  L \otimes_k \mc{I}_k\left( \bigcup_{x \in X} \{x\} \right)
     =  L \otimes_k \left( \bigcap_{x \in X} \mc{I}_k(\{x\}) \right)  \\
    &=  \bigcap_{x \in X} \left( L \otimes_k \mc{I}_k(\{x\}) \right)
     =  \bigcap_{x \in X} \mc{I}_k(\{x\})_L
     =  \bigcap_{x \in X} \mc{I}_L(\{x\})
     =  \mc{I}_L(X) \,.
  \end{align*}
  This proves the claim.
\end{proof}


\begin{lemma}[Transitivity of Zariski density]
  Let $V$ be a finite-dimensional $k$-vector space and let $X \subseteq Y \subseteq Z \subseteq V$.
  If $X$ is Zariski-dense in $Y$ and $Y$ is Zariski-dense in $Z$, then $X$ is Zariski-dense in $Z$.
\end{lemma}


\begin{proof}
  It follows for every polynomial function $f \colon V \to k$ from $\restrict{f}{X} = 0$ that $\restrict{f}{Y} = 0$ and thus $\restrict{f}{Z} = 0$.
\end{proof}


\begin{corollary}
  \label{corollary: Zariski dense scalar extension}
  Let $V$ be a finite-dimensional $k$-vector space and let
  \[
              X
    \subseteq Y
    \subseteq V
    \subseteq V_L \,.
  \]
  If $X$ is Zariski dense in $Y$ over $k$ then $X$ is also Zariski dense in $Y$ over $L$.
\end{corollary}


\begin{proof}
  We have that $\mc{I}_k(X) = \mc{I}_k(Y)$ because $X$ is Zariski dense in $Y$ over $k$.
  It follows that
  \[
      \mc{I}_L(X)
    = \mc{I}_k(X)_L
    = \mc{I}_k(Y)_L
    = \mc{I}_L(Y) \,,
  \]
  which shows that $X$ is Zariski dense in $Y$ over $L$.
\end{proof}





% We can generalize this obeservation in the language of extension of scalars:
% 
% 
% \begin{lemma}\label{lemma: W Zariski dense in W_L}
%   Let $V$ be a finite-dimensional $k$-vector space and let $L/k$ be a field extension.
%   Then $V$ is Zariski dense in $V_L$ over $L$.
% \end{lemma}
% \begin{proof}
%   Let $w_1, \dotsc, w_n$ be a $k$-basis of $W$.
%   Then $1 \otimes w_1, \dotsc, 1 \otimes w_n$ is an $L$-basis of $W_L$.
%   Using this bases we have an isomorphism of $k$-vector spaces
%   \[
%             \phi
%     \colon  k^n \to W,
%     \quad   e_i
%     \mapsto w_i
%   \]
%   and an isomorphism of $L$-vector spaces
%   \[
%             \psi
%     \colon  L^n
%     \to     W_L,
%     \quad   e_i
%     \mapsto 1 \otimes w_i
%   \]
%   and the following commutative diagram:
%   \[
%     \begin{tikzcd}
%         W
%         \arrow[equal]{r}{\sim}
%         \arrow[hook]{d}
%       & k^n
%         \arrow[hook]{d}
%       \\
%         W_L
%         \arrow[equal]{r}{\sim}
%       & L^n
%     \end{tikzcd}
%   \]
%   The isomorphism $\psi$ of $L$-vector spaces induces the isomorphism
%   \[
%             \psi^*
%     \colon  \mc{P}_L(W_L)
%     \to     \mc{P}_L(L^n),
%     \quad   h
%     \mapsto h \circ \psi
%   \]
%   of $L$-algebras.
%   For $f \in \mc{P}_L(W_L)$ with $f_{|W} = 0$ we have $g \coloneqq \psi^*(f) \in \mc{P}_L(L^n)$ with $g_{|k^n} = 0$.
%   Since $k^n$ is Zariski dense in $L^n$ over $L$ we find that $g = 0$.
%   Since $\psi^*$ is an isomorphism of $L$-algebras it follows that $f = 0$.
% \end{proof}
% 
% 
% As we have seen in the last chapter we have $\mc{P}_k(W)_L \cong \mc{P}_L(W_L)$ as $L$-algebras.
% The problem of the constructed isomorphism is that it depends on choosing a $k$-basis of $W$ and an $L$-basis of $W_L$.
% We will now construct an isomorphism which does not depend on such choice.
% 
% 
% \begin{lemma}
%   Let $W$ be a finite-dimensional $k$-vector space. Then there exists an unique map
%   \[
%             \iota
%     \colon  \mc{P}_k(W)
%     \to     \mc{P}_L(W_L)
%   \]
%   such that for every $f \in \mc{P}_k(W)$ the diagram
%   \[
%     \begin{tikzcd}[sep = large]
%         W
%         \arrow[hook]{r}{\can_W}
%         \arrow[swap]{d}{f}
%       & W_L
%         \arrow{d}{\iota(f)}
%       \\
%         k
%         \arrow[hook]{r}
%       & L
%     \end{tikzcd}
%   \]
%   commutes.
%   $\iota$ is a monomorphism of $k$-algebras.
% \end{lemma}
% \begin{proof}
%   Let $w_1, \dotsc, w_n$ be a $k$-basis of $W$ and
%   \[
%         \varphi_1, \dotsc, \varphi_n
%     \in \mc{P}_k(W)
%   \]
%   the corresponding coordinate functions. Also let
%   \[
%         \psi_1, \dotsc, \psi_n
%     \in \mc{P}_L(W_L)
%   \]
%   be the coordinate functions of the $L$-basis $1 \otimes w_1, \dotsc, 1 \otimes w_n$ of $W_L$.
%   Then we define $\iota$ to be the unique homomorphism of $k$-algebras with
%   \[
%       \iota(\varphi_i)
%     = \psi_i
%     \text{ for all }
%     1 \leq i \leq n \,.
%   \]
%   Since $\psi_1, \dotsc, \psi_n$ are algebraically independent over $L$ we know that $\iota$ is injective.
%   We have for every $f \in k[X_1, \dotsc, X_n]$ and $w = \sum_{i=1}^n \mu_i w_i \in W$
%   \begin{align*}
%      &\,  \iota(f(\varphi_1, \dotsc, \varphi_n))(\can_W(w)) \\
%     =&\,  f(\psi_1, \dotsc, \psi_n)\left( \sum_{i=1}^n \mu_i (1 \otimes w_i) \right) \\
%     =&\,  f(\mu_1, \dotsc, \mu_n)
%      =    f(\varphi_1, \dotsc, \varphi_n)\left( \sum_{i=1}^n \mu_i w_i \right) \\
%     =&\,  f(\varphi_1, \dotsc, \varphi_n)(w) \,.
%   \end{align*}
%   Since $\varphi_1, \dotsc, \varphi_n$ generate $\mc{P}_k(W)$ as a $k$-algebra it follows that the diagram does commutes.
%   
%   All that’s left to show is the uniqueness of $\iota$.
%   For this we will use our newly acquired wisdom about Zariski density.
%   Suppose we have a map
%   \[
%             i
%     \colon  \mc{P}_k(W)
%     \to     \mc{P}_L(W_L)
%   \]
%   such that the diagram
%   \[
%     \begin{tikzcd}[sep = large]
%         W
%         \arrow[hook]{r}{\can_W}
%         \arrow[swap]{d}{f}
%       & W_L
%         \arrow{d}{\iota(f)}
%       \\
%         k
%         \arrow[hook]{r}
%       & L
%     \end{tikzcd}
%   \]
%   commutes for all $f \in \mc{P}_k(W)$.
%   For every polynomial map $f \in \mc{P}_k(W)$ we then have \mbox{$\iota(f)-i(f) \in \mc{P}_L(W_L)$} with
%   \[
%       \left(
%         \iota(f)-i(f)
%       \right)_{|W}
%     = 0 \,.
%   \]
%   Since $W$ is Zariski dense in $W_L$ over $L$ it follows that $\iota(f) - i(f) = 0$ and therefore $\iota(f) = i(f)$.
% \end{proof}
% 
% 
% This inclusion gives us the desired isomorphism of $L$-algebras
% \[
%         \mc{P}_k(W)_L
%   \cong \mc{P}_L(W_L)
% \]
% for free.
% 
% 
% \begin{proposition}
%   Let $k$ be a finite-dimensional $k$-vector space.
%   Then the map
%   \[
%             \Phi
%     \colon  \mc{P}_k(W)_L
%     \to     \mc{P}_L(W_L),
%     \quad   \lambda \otimes f
%     \mapsto \lambda \iota(f)
%   \]
%   is an isomorphism of $L$-algebras, where $\iota \colon \mc{P}_k(W) \to \mc{P}_L(W_L)$ is the inclusion as before.
% \end{proposition}
% \begin{proof}
%   Let $v_1, \dotsc, v_n$ be a $k$-basis of $W$ and $\varphi_1, \dotsc, \varphi_n \in \mc{P}_k(W)$ the corresponding coordinate functions.
%   Then $\varphi^\alpha = \varphi_1^{\alpha_1} \dotsm \varphi_n^{\alpha_n}$ with $\alpha \in \Natural^n$ are a $k$-basis of $\mc{P}_k(W)$ and an $L$-basis of $\mc{P}_L(W_L)$ (where we identify $W$ with the corresponding $k$-vector subspace of $W_L$ and $\mc{P}_k(W)$ with the corresponding $k$-subalgebra of $\mc{P}_L(W_L)$ under $\iota$).
%   Therefore the statement follows from corollary \ref{corollary: inclusion to bijection algebras}.
% \end{proof}
% 
% 
% Using this isomorphism we will identify $\mc{P}_k(W)_L$ with $\mc{P}_L(W_L)$.
% We also identify $\mc{P}_k(W)$ with its image under $\iota$.
% 
% 
% We can also combine all these commuting diagrams into a big one:
% Given a $k$-vector space $W$ and a $k$-basis $w_1, \dotsc, w_n$ of $W$ we have an isomorphism
% \[
%         k[X_1, \dotsc, X_n]
%   \cong \mc{P}_k(W)
% \]
% of $k$-vector spaces, and corresponding with the $L$-basis $1 \otimes w_1, \dotsc, 1 \otimes w_n$ of $W_L$ we have an isomorphism of $L$-algebras
% \[
%         L[X_1, \dotsc, X_n]
%   \cong \mc{P}_L(W_L) \,.
% \]
% This gives us the following commutative diagram.
% \[
%   \begin{tikzcd}[sep = large]
%       \mc{P}_k(W)
%       \arrow[hook]{r}{\iota}
%       \arrow[equal, swap]{d}{\wr}
%     & \mc{P}_L(W_L)
%       \arrow[equal]{d}{\wr}
%     \\
%       k[X_1, \dotsc, X_n]
%       \arrow[hook]{r}
%     & L[X_1, \dotsc, X_n]
%   \end{tikzcd}
% \]
% Given the inclusion $\iota \colon \mc{P}_k(W) \to \mc{P}_L(W_L)$ and the corresponding isomorphism of $L$-algebras $\mc{P}_k(W)_L \cong \mc{P}_L(W_L)$ we also have the following commutative diagram.
% \[
%   \begin{tikzcd}[sep = large]
%       \mathcal{P}_k(W)
%       \arrow[hook]{r}
%       \arrow[hook, swap]{d}{\iota}
%     & \mathcal{P}_L(W_L)
%     \\
%       \mathcal{P}_k(W)_L
%       \arrow[equal]{ru}[rotate=30]{\sim}
%     & {}
%   \end{tikzcd}
% \]
% We get a similar diagram for polynomial rings.
% (See the appendix \ref{app: extension of scalars} for more details.)
% \[
%   \begin{tikzcd}[sep = large]
%       k[X_1, \dotsc, X_n]
%       \arrow[hook]{r}
%       \arrow[hook]{d}
%     & L[X_1, \dotsc, X_n]
%     \\
%       k[X_1, \dotsc, X_n]_L
%       \arrow[equal]{ru}[rotate=30]{\sim}
%     & {}
%   \end{tikzcd}
% \]
% The isomorphism $\mc{P}_k(W) \cong k[X_1, \dotsc, X_n]$ of $k$-algebras also induces an isomorphism $\mc{P}_k(W)_L \cong k[X_1, \dotsc, X_n]_L$ of $L$-algebras which results in the following commutative diagram.
% \[
%   \begin{tikzcd}[sep = large]
%       \mc{P}_k(W)_L
%       \arrow[equal]{d}{\wr}
%     & \mc{P}_k(W)
%       \arrow[left hook ->]{l}
%       \arrow[equal]{d}{\wr}
%     \\
%       k[X_1, \dotsc, X_n]_L
%     & k[X_1, \dotsc, X_n]
%       \arrow[left hook ->]{l}
%   \end{tikzcd}
% \]
% We also have the following commutative diagram of $L$-algebras and isomorphisms of such.
% \[
%   \begin{tikzcd}[sep = large]
%       \mc{P}_k(W)_L
%       \arrow[equal]{d}{\wr}
%       \arrow[equal]{r}{\sim}
%     & \mc{P}_L(W_L)
%       \arrow[equal]{d}{\wr}
%     \\
%       k[X_1, \dotsc, X_n]_L
%       \arrow[equal]{r}{\sim}
%     & L[X_1, \dotsc, X_n]
%   \end{tikzcd}
% \]
% By putting all of this together we get the following commutative diagram:
% \[
%   \begin{tikzcd}
%       {}
%     & \mc{P}_k(W)
%       \arrow[left hook ->]{dl}
%       \arrow[hook]{dr}
%       \arrow[equal, near start]{dd}{\wr}
%     & {}
%     \\
%       \mc{P}_k(W)_L
%       \arrow[equal, near start]{rr}{\sim}
%       \arrow[equal]{dd}{\wr}
%     & {}
%     & \mc{P}_L(W_L)
%       \arrow[equal]{dd}{\wr}
%     \\
%       {}
%     & k[X_1, \dotsc, X_n]
%       \arrow[left hook ->]{dl}
%       \arrow[hook]{dr}
%     & {}
%     \\
%       k[X_1, \dotsc, X_n]_L
%       \arrow[equal]{rr}{\sim}
%     & {}
%     & L[X_1, \dotsc, X_n]
%   \end{tikzcd}
% \]

\begin{fluff}
  Let $V$ be a finite-dimensional $k$-vector space.
  To prove the last part of the following proposition we need to slightly generalize our notion of Zariski density:
  
  For a subset $X \subseteq V$ a function $X \to k$ is \emph{polynomial} if it is the restriction of a polynomial function $V \to k$.
  If $X$ is Zariski dense in $Y \subseteq V$ and $f \colon Y \to k$ is a polynomial function with $\restrict{f}{X} = 0$ it then follows that $f = 0$.
  
  For a subset $X \subseteq V$ a function $f \colon X \to k$ is \emph{rational} if there exists polynomial functions $g, h \colon X \to k$ with $h(x) \neq 0$ for every $x \in X$ and
  \[
      f(x)
    = \frac{g(x)}{h(x)}
  \]
  for every $x \in X$.
  Note that a rational function $f = g/h$ vanishes if and only if its numerator $g$ vanishes.
  If $X$ is Zariski dense in $Y \subseteq V$ and $f \colon Y \to k$ is rational with $\restrict{f}{X} = 0$, it thus follows that $f = 0$.
  
  Both the polynomial functions $X \to k$ and rational functions $X \to k$ form $k$-algebras via pointwise addition and scalar multiplication of functions.
\end{fluff}


\begin{warning}
  While every polynomial function $f \colon X \to k$ can be extended to a polynomial function $V \to k$, the same does not hold for rational functions.
  Consider for example $\GL_n(k) \subseteq \Mat_n(k)$ and the rational function $f \colon \GL_n(k) \to k$ given by
  \[
              f(A)
    \defined  \frac{1}{\det(A)} \,.
  \]
  Suppose that $f$ could be extended to a rational function $\hat{f} \colon \Mat_n(k) \to k$.
  Then the function $g \colon \Mat_n(k) \to k$ given by $g(A) = \hat{f}(A) \det(A) - 1$ is also rational, and $\restrict{g}{\GL_n(k)} = 0$.
  It follows that $g = 0$ because $\GL_n(k)$ is Zariski dense in $\Mat_n(k)$, and therefore that $\hat{f}(A) \det(A) = 1$ for all $A \in \Mat_n(K)$.
  But this is not possible.
  (This shows more generally that $f$ cannot be extended to any subset $X \subseteq \Mat_n(k)$ with $\GL_n(k) \subsetneq X$ as a rational function.)
\end{warning}


\begin{proposition}
  \leavevmode
  If we consider the groups $\GL_n(k), \GL_n(L), \SL_n(k), \SL_n(L)$ as subsets of $\Mat_n(L)$ then
  \begin{enumerate}
    \item
      $\GL_n(k)$ is Zariski dense in $\Mat_n(L)$ over $L$,
    \item
      $\GL_n(k)$ is Zariski dense in $\GL_n(L)$ over $L$,
    \item
      $\SL_n(k)$ is Zariski dense in $\SL_n(L)$ over $L$.
  \end{enumerate}
\end{proposition}
\begin{proof}
  \leavevmode
  \begin{enumerate}
    \item
      We have already seen that $\GL_n(k)$ is Zariski dense in $\Mat_n(k)$ over $k$.
      It follows that $\GL_n(k)$ is Zariski dense in $\Mat_n(k)$ over $L$ by Corollary~\ref{corollary: Zariski dense scalar extension}, with $\Mat_n(k)$ being Zariski dense in $\Mat_n(L)$ over $L$ by Corollary~\ref{corollary: kn is Zariski dense in Ln}.
      It follows from the transitivity of Zariski density that $\GL_n(k)$ is Zariski dense in $\Mat_n(L)$ over $L$.
    \item
      This follows from a).
    \item
      Let $f \colon \Mat_n(L) \to L$ be a polynomial function with $\restrict{f}{\SL_n(k)} = 0$.
      Consider the map
      \[
                p
        \colon  \GL_n(L)
        \to     \SL_n(L) \,,
        \quad   A
        \mapsto \begin{pmatrix}
                  \det(A)^{-1}  &   &         &     \\
                                & 1 &         &     \\
                                &   & \ddots  &     \\
                                &   &         & 1
                \end{pmatrix}
                \cdot
                A \,,
      \]
      which divides the first row of $A$ by $\det(A)$.
      For every $A \in \SL_n(L)$ we have that $p(A) = A$, so we may think of $p$ as a (rational) retraction of $\GL_n(L)$ onto $\SL_n(L)$.
      Note that $p$ restricts to a map $\restrict{p}{\GL_n(k)} \colon \GL_n(k) \to \SL_n(k)$.
      
      The map $p$ is rational in every coordinate, so it follows that the composition $\restrict{f}{\SL_n(L)} \circ p \colon \GL_n(L) \to L$ is rational.
      For every $A \in \GL_n(k)$ we have that $p(A) \in \SL_n(k)$ and therefore
      \[
          (\restrict{f}{\SL_n(L)} \circ p)(A)
        = f(p(A))
        = 0 \,.
      \]
      It follows that $(\restrict{f}{\SL_n(L)} \circ p)(A) = 0$ for every $A \in \GL_n(L)$ because $\GL_n(k)$ is Zariski-dense in $\GL_n(L)$ over $L$.
      It then follows for every $A \in \SL_n(L)$ that
      \[
          0
        = (\restrict{f}{\SL_n(L)} \circ p)(A)
        = f(p(A))
        = f(A) \,.
      \]
      This shows that $\restrict{f}{\SL_n(L)} = 0$.
    \qedhere
  \end{enumerate}
\end{proof}


\begin{proposition}
  Let $k$ be an infinite field and let $\GL_n(k)$ act on $\Mat_n(k)$ by conjugation.
  Let $f \colon \Mat_n(k) \to k$ be a $\GL_n(k)$-invariant polynomial function and let $D \subseteq \Mat_n(k)$ be the subset of diagonal matrices.
  If $\restrict{f}{D} = 0$ then $f = 0$.
\end{proposition}
\begin{proof}
  Let $L$ be an algebraic closure of $k$ and let $\GL_n(L)$ acts on $\Mat_n(L)$ by conjugation.
  Let $\overline{f} \colon \Mat_n(L) \to L$ be the unique $L$-polynomial extension of $f$ (as described in Proposition~\ref{proposition: consistent choice of identification for polynomial functions}).
  
  The map $\overline{f}$ is $\GL_n(L)$-invariant:
  We have for all $S \in \GL_n(k)$, $A \in \Mat_n(k)$ that
  \[
      \overline{f}(SA)
    = f(SA)
    = f\left( SASS^{-1} \right)
    = f(S.(AS))
    = f(AS)
    = \overline{f}(AS)  \,,
  \]
  and therefore $\Phi(S,A) = 0$ for the map $\Phi \colon \Mat_n(L) \times \Mat_n(L) \to L$ given by
  \[
      \Phi(A,B)
    = \overline{f}(AB) - \overline{f}(BA)
  \]
  for all $A, B \in \Mat_n(L)$.
  Both $\GL_n(k) \subseteq \GL_n(L)$ and $\Mat_n(k) \subseteq \Mat_n(L)$ are Zariski dense over $L$, so it follows from Lemma~\ref{lemma: product of Zariski dense is again Zariski dense} that $\GL_n(k) \times \Mat_n(k) \subseteq \GL_n(L) \times \Mat_n(L)$ is Zariski dense over $L$.
  It thus follows from $\Phi_{\GL_n(k) \times \Mat_n(k)} = 0$ that $\Phi_{\GL_n(L) \times \Mat_n(L)} = 0$.
  For all $S \in \GL_n(L)$, $A \in \Mat_n(L)$ we thus have that
  \[
      \overline{f}(S.A)
    = \overline{f}(SAS^{-1})
    = \overline{f}(A S^{-1} S)
    = \overline{f}(A).
  \]
  This shows that $\overline{f}$ is $\GL_n(L)$-invariant.
    
  Let $D_k \subseteq \Mat_n(k)$ and $D_L \subseteq \Mat_n(L)$ be the subsets of diagonal matrices.
  It follows from $\restrict{f}{D_k} = 0$ that $\restrict{\overline{f}}{D_k} = 0$, from which it further follows that $\restrict{\overline{f}}{D_L} = 0$ because $D_k$ is Zariski-dense in $D_L$ over $L$ by Corollary~\ref{corollary: kn is Zariski dense in Ln}.
  It follows by Corollary~\ref{corollary: diagonal matrices dense alg closed} that $\overline{f} = 0$ and therefore that $f = 0$.
\end{proof}


\noindent\hrulefill \, Current progress of reworking these notes. \hrulefill




