\section{The Nullstellensätze}


\begin{fluff}
  We can associate to every subset $X \subseteq V$ its vanishing ideal $\mc{I}(X) \idealleq \mc{P}(V)$, and to every ideal $I \idealleq \mc{P}(V)$ its vanishing set $\mc{V}(I) \subseteq V$, resulting in maps $\mc{I}, \mc{V}$ as follows:
  \[
    \begin{tikzcd}
        \{ \text{subsets $X \subseteq V$} \}
        \arrow[shift left]{r}[above]{\mc{I}}
      & \{ \text{ideals $I \idealleq \mc{P}(V)$} \}
        \arrow[shift left]{l}[below]{\mc{V}}
    \end{tikzcd}
  \]
  In general the maps $\mc{I}, \mc{V}$ will neither be injective nor surjective.
  But we will see in this section that when we restrict our attention to suitable classes (or rather sets) of subsets $X \subseteq V$ and ideals $I \idealleq \mc{P}(V)$ the maps $\mc{V}$ and $\mc{I}$ not only restrict to bijections, but that they become inverse to each other.
\end{fluff}


\begin{lemma}
  \label{lemma: galois connection for vanishing ideals and zero sets}
  For every subset $X \subseteq V$ and ideal $I \idealleq \mc{P}(V)$ we have that
  \[
          X \subseteq \mc{V}(I)
    \iff  \mc{I}(X) \supseteq I \,.
  \]
\end{lemma}


\begin{proof}
  Both conditions state that $f(x) = 0$ for all $f \in I$, $x \in X$.
\end{proof}


\begin{lemma}
  \label{lemma: properties of V and I}
  Let $X \subseteq V$ be a subset and let $I \idealleq \mc{P}(V)$ be an ideal.
  Then
  \begin{enumerate}
    \item
      \label{enumerate: VI is monotone}
      $X \subseteq \mc{V}(\mc{I}(X))$,
    \item
      \label{enumerate: IV is monotone}
      $I \subseteq \mc{I}(\mc{V}(I))$,
    \item
      \label{enumerate: IVI = I}
      $\mc{I}(\mc{V}(\mc{I}(X))) = \mc{I}(X)$,
    \item
      \label{enumerate: VIV = V}
      $\mc{V}(\mc{I}(\mc{V}(I))) = \mc{V}(I)$.
  \end{enumerate}
\end{lemma}


\begin{proof}
  \leavevmode
  \begin{enumerate}
    \item
      This is by Lemma~\ref{lemma: galois connection for vanishing ideals and zero sets} equivalent to $\mc{I}(X) \supseteq \mc{I}(X)$ .
    \item
      This is by Lemma~\ref{lemma: galois connection for vanishing ideals and zero sets} equivalent to $\mc{V}(I) \subseteq \mc{V}(I)$.
    \item
      It follows from $X \subseteq \mc{V}(\mc{I}(X))$ that $\mc{I}(X) \supseteq \mc{I}(\mc{V}(\mc{I}(X)))$ because $\mc{I}$ is order-reversing.
      That $\mc{I}(X) \subseteq \mc{I}(\mc{V}(\mc{I}(X)))$ is equivalent to $\mc{V}(\mc{I}(X)) \supseteq \mc{V}(\mc{I}(X)) $ by Lemma~\ref{lemma: galois connection for vanishing ideals and zero sets}.
%       That $\mc{I}(X) \subseteq \mc{I}(\mc{V}(\mc{I}(X)))$ follows from part~\ref*{enumerate: IV is monotone}.
%       The inclusion $\mc{I}(X) \supseteq \mc{I}(\mc{V}(\mc{I}(X)))$ is by Lemma~\ref{lemma: galois connection for vanishing ideals and zero sets} equivalent to $X \subseteq \mc{V}(\mc{I}(\mc{V}(\mc{I}(X))))$, which follows from part~\ref*{enumerate: VI is monotone} because
%       \[
%                   X
%         \subseteq \mc{V}(\mc{I}(X))
%         \subseteq \mc{V}(\mc{I}(\mc{V}(\mc{I}(X)))) \,.
%       \]
    \item
      It follows from $I \subseteq \mc{I}(\mc{V}(I))$ that $\mc{V}(I) \supseteq \mc{V}(\mc{I}(\mc{V}(I)))$ because $\mc{V}$ is order-reversing.
      That $\mc{V}(I) \subseteq \mc{V}(\mc{I}(\mc{V}(I)))$ is equivalent to $\mc{I}(\mc{V}(I)) \supseteq \mc{I}(\mc{V}(I)) $ by Lemma~\ref{lemma: galois connection for vanishing ideals and zero sets}.
    \qedhere
  \end{enumerate}
\end{proof}


\begin{definition}
  An ideal $I \idealleq \mc{P}(V)$ is a \emph{vanishing ideal} if $I = \mc{I}(X)$ for some $X \subseteq V$.
\end{definition}


\begin{corollary}
  \label{corollary: bijection induced by Galois correspondence}
  The maps $\mc{I}$, $\mc{V}$ restrict to the following mutually inverse bijections:
  \[
    \begin{matrix}
        \left\{
          \begin{tabular}{c}
              algebraic subsets \\
              $X \subseteq V$
          \end{tabular}
        \right\}
      & \begin{tikzcd}[column sep = large]
            {}
            \arrow[shift left]{r}{\mc{I}}
          & {}
            \arrow[shift left]{l}{\mc{V}}
        \end{tikzcd}
      & \left\{
          \begin{tabular}{c}
            vanishing ideals \\
            $I \idealleq \mc{P}(V)$
          \end{tabular}
        \right\}
    \end{matrix}
  \]
  i.e.\ for every algebraic subset $X \subseteq V$ we have that
  \[
    \mc{V}(\mc{I}(X)) = X \,,
  \]
  and for every vanishing ideal $I \idealleq \mc{P}(V)$ we have that
  \[
    \mc{I}(\mc{V}(I)) = I \,.
  \]
\end{corollary}


\begin{proof}
  The two identities are just reformulations of parts \ref*{enumerate: IVI = I}, \ref*{enumerate: VIV = V} of Lemma~\ref{lemma: properties of V and I}.
\end{proof}


\begin{lemma}
  \label{lemma: correspence between points and vanishing maximal ideals}
  An ideal $\mf{m} \idealleq \mc{P}(V)$ is of the form $\mf{m} = \mf{m}_a$ for some $a \in V$ if and only if it is both a maximal ideal and a vanishing ideal.
\end{lemma}


\begin{proof}
  The ideal $\mf{m}_a$ is maximal and it is a vanishing ideal because $\mf{m} = \mc{I}(a)$.
  
  Suppose on the other hand that $\mf{m}$ is both a maximal ideal and a vanishing ideal.
  Then $\mc{I}(\mc{V}(\mf{m})) = \mf{m}$ by Corollary~\ref{corollary: bijection induced by Galois correspondence} because $\mf{m}$ is a vanishing ideal.
  It then follows that $\mc{V}(\mf{m}) \neq \emptyset$ because otherwise $\mf{m} = \mc{I}(\emptyset) = \mc{P}(V)$, which contradicts $\mf{m}$ being a proper ideal.
  It follows that there exists some $a \in \mc{V}(\mf{m})$.
  Then
  \[
              \mf{m}
    =         \mc{I}(\mc{V}(\mf{m}))
    \subseteq \mc{I}(a)
    =         \mf{m}_a
  \]
  and it follows that $\mf{m} = \mf{m}_a$ because both $\mf{m}$ and $\mf{m}_a$ are maximal.
\end{proof}

\begin{corollary}
  \label{corollary: general correspondence for algebraic subsets and points}
  The maps $\mc{I}, \mc{V}$ restrict to the following mutually inverse bijections:
  \[
    \begin{matrix}
        \left\{
          \begin{tabular}{c}
              algebraic subsets \\
              $X \subseteq V$
          \end{tabular}
        \right\}
      & \begin{tikzcd}[column sep = large]
            {}
            \arrow[shift left]{r}{\mc{I}}
          & {}
            \arrow[shift left]{l}{\mc{V}}
        \end{tikzcd}
      & \left\{
          \begin{tabular}{c}
            vanishing ideals \\
            $I \idealleq \mc{P}(V)$
          \end{tabular}
        \right\}
      \\
        {}
      & {}
      & {}
      \\
        \rotatebox[origin=c]{90}{$\subseteq$}
      & {}
      & \rotatebox[origin=c]{90}{$\subseteq$}
      \\
        {}
      & {}
      & {}
      \\
        \left\{
          \begin{tabular}{c}
            points $a \in V$
          \end{tabular}
        \right\}
      & \begin{tikzcd}[column sep = large]
            {}
            \arrow[shift left]{r}{\mc{I}}
          & {}
            \arrow[shift left]{l}{\mc{V}}
        \end{tikzcd}
      & \left\{
          \begin{tabular}{c}
            vanishing ideals \\
            $\mf{m} \idealleq \mc{P}(V)$ \\
            which are maximal
          \end{tabular}
        \right\}
    \end{matrix}
  \]
\end{corollary}


\begin{proof}
  This follows from Corollary~\ref{corollary: bijection induced by Galois correspondence} because by Lemma~\ref{lemma: correspence between points and vanishing maximal ideals} the vanishing ideals which correspond to points $a \in V$ are precisely the vanishing ideals which are also maximal.
\end{proof}


\begin{remark}[Galois connections]
  \label{remark: galois connection}
  We will use this opportunity to introduce the notion of a \emph{Galois connections}, although we will not use it in the rest of the text.
  Let $(A,\leq)$, $(B, \leq)$ be two partially ordered sets.
  \begin{enumerate}
    \item
      An \emph{antitone Galois connection} consists of two functions $f \colon A \to B$, $g \colon B \to A$ which are order-reversing, i.e.\ $f(a) \geq f(a')$ for all $a \leq a'$ and $g(b) \geq g(b')$ for all $b \leq b'$, such that
      \begin{equation}
        \label{equation: antitone galois connection}
              a \leq g(b)
        \iff  f(a) \geq b
      \end{equation}
      for all $a \in A$, $b \in B$;
      the condition~\eqref{equation: antitone galois connection} may also be expressed as
      \[
              a \leq g(b)
        \iff  b \leq f(a) \,.
      \]
      Lemma~\ref{lemma: galois connection for vanishing ideals and zero sets} states that $\mc{I}, \mc{V}$ form a Galois connection (between suitable sets).
      Lemma~\ref{lemma: properties of V and I} generalizes for an arbitrary Galois connection to
      \begin{enumerate}[label = \alph*')]
        \item
          \label{enumerate: gf is monotone}
          $a \leq g(f(a))$ for every $a \in A$,
        \item
          \label{enumerate: fg is monotone}
          $b \leq f(g(b))$ for every $b \in B$,
        \item
          \label{enumerate: fgf = f}
          $f(g(f(a))) = f(a)$ for every $a \in A$,
        \item
          \label{enumerate: gfg = g}
          $g(f(g(b))) = g(b)$ for every $b \in B$,
      \end{enumerate}
      and Corollary~\ref{corollary: bijection induced by Galois correspondence} generalizes to $f, g$ restricting to mutually inverse order-re\-ver\-sing bijections between the sets
      \[
        \{ f(a) \suchthat a \in A \}
        \qquad\text{and}\qquad
        \{ g(b) \suchthat b \in B \} \,.
      \]
      Note that the roles of $f$ and $g$ are symmetric, in the sense that $((A, \leq), (B, \leq), f, g)$ is an antitone Galois connection if and only if $((B, \leq), (A, \leq), g, f)$ is an antitone Galois connection.
      One may visualize an antitone Galois-connection as follows:
      \[
        \begin{tikzcd}
            (A,\leq)
            \arrow[shift left]{r}[above]{f}
          & (B,\leq)
            \arrow[shift left]{l}[below]{g}
        \end{tikzcd}
      \]
    \item
      Instead of \emph{antitone} Galois connections, one can also consider \emph{monotone} Galois connections.
      The conditition of $f, g$ being order-reversing is then replaced by the requirement of $f, g$ being order-preserving, i.e.\ monotone, and the conditition~\eqref{equation: antitone galois connection} is adjusted to
      \[
              a \leq g(b)
        \iff  f(a) \leq b \,.
      \]
      Note that this requirement is not symmetric in $f$ and $g$:
      If $((A, \leq), (B, \leq), f, g)$ is a monotone Galois connection then $((B, \leq), (A, \leq), g, f)$ is not necessarily a monotone Galois connection.
      This non-symmetry is reflected in the fact that $f$ is referred to as the \emph{left adjoint} and $g$ is referred to as the \emph{right adjoint} of this (monotone) Galois connection.
      The consequences \ref*{enumerate: gf is monotone}, \ref*{enumerate: fgf = f}, \ref*{enumerate: gfg = g} still hold, but \ref*{enumerate: fg is monotone} has to be replaced by
      \[
        f(g(b)) \leq b \,.
      \]
      The maps $f, g$ still restrict to mutually inverse bijections as above.
      One may visualize a monotone Galois connection with left adjoint $f$ and right adjoint $g$ as follows:
      \[
        \begin{tikzcd}
            (A,\leq)
            \arrow[bend left]{r}[above]{f}
            \arrow[phantom]{r}[rotate=90]{\vdash}
          & (B,\leq)
            \arrow[bend left]{l}[below]{g}
        \end{tikzcd}
      \]
    \item
      We have the following connections between antitone and monoton Galois connections:
      \[
          \begin{tikzcd}
              (A,\leq)
              \arrow[bend left]{r}[above]{f}
              \arrow[phantom]{r}[rotate=90]{\vdash}
            & (B,\leq)
              \arrow[bend left]{l}[below]{g}
          \end{tikzcd}
          \quad\iff\quad
          \begin{tikzcd}
              (A,\leq)
              \arrow[shift left]{r}[above]{f}
            & (B,\leq^{\op})
              \arrow[shift left]{l}[below]{g}
          \end{tikzcd}
      \]
      Here $\leq^{\op}$ denotes the partial order given by
      \[
              x \leq^{\op} y
        \iff  x \geq y \,.
      \]
    \item
      One can think about the partially ordered sets $(A, \leq)$ and $(B, \leq)$ as categories $\mc{A}, \mc{B}$ in the usual way, i.e.\ the objects of $\mc{A}$ (resp.\ $\mc{B}$) are the elements of $A$ (resp.\ of $B$) and for all $x, y \in \mc{A}$ (resp.\ $x, y \in \mc{B}$) there exists a unique morphism $x \to y$ if $x \leq y$, and no morphism $x \to y$ otherwise.
      
      A pair of order-preserving maps $f \colon A \to B$, $g \colon B \to A$ can be regarded as a pair of (covariant) functors $F \colon \mc{A} \to \mc{B}$ and $G \colon \mc{B} \to \mc{A}$, whose action on objects are given by the maps $f, g$ and whose action on morphisms is the only possible one.
      Then $f, g$ form a monotone Galois connection with $f$ left adjoint to $g$ if and only if the functor $F$ is left adjoint to the functor $G$.
      
      A pair of order-reversing maps $f \colon A \to B$, $g \colon B \to A$ can then be regarded as a pair of contravariant functors $F \colon \mc{A} \to \mc{B}$ and $G \colon \mc{B} \to \mc{A}$.
      Then $f, g$ define an antitone Galois connection if the functors $F, G$ are adjoint on the right.
      
      This abstract viewpoint has some direct consequences for Galois connections:
      \begin{itemize}
        \item
          Left (resp.\ right) adjoints are unique up to isomorphism.
          The only isomorphisms in $\mc{A}, \mc{B}$ are the identies (by the anti-symmetry of $\leq$) so it follows that $F$ is uniquely determined by $G$ and vice versa.
          It follows that in a Galois connection $f,g$ the map $f$ is uniquely determined by $g$ and vice versa.
          
        \item
          If $f, g$ form a monotone adjoint Galois connection with $f$ left adjoint to $g$ then $F$ is left adjoint to $G$ and it follows that $F$ preserves colimits and $G$ preserves limits.
          It follows in particular that $F$ preserves coproducts while $G$ preserves products.
          Coproducts in $\mc{A}$ are just suprema in $(A, \leq)$, and products in $\mc{B}$ are just infima in $(B, \leq)$.
          It thus follows that $f$ preserves suprema while $g$ preserves infima.
          
          If $f,g$ form an antitone Galois connection between $(A, \leq)$ and $(B, \leq)$ then $f,g$ form a monotone Galois connection between $(A,\leq)$ and $(B,\leq^\op)$.
          It then follows from the above that both $f$ and $g$ turn suprema into infima.
          
          For the maps $\mc{I}, \mc{V}$ this means that
          \[
              \mc{I}\left( \bigcup_{i \in I} X_i \right)
            = \bigcap_{i \in I} \mc{I}(X_i)
          \]
          for every family $(X_i)_{i \in I}$ of subsets $X_i \subseteq V$, which we have already seen in Lemma~\ref{lemma: basic properties of I}, and that
          \[
              \mc{V}\left( \sum_{j \in J} I_j \right)
            = \bigcap_{i \in I} \mc{V}(I_j)
          \]
          for every family $(I_j)_{j \in J}$ of ideal $I_j \idealleq \mc{P}(V)$, which we will see in Lemma~\ref{lemma: intersections and unions of Zariski closed sets}.
      \end{itemize}
  \end{enumerate}
\end{remark}


\begin{fluff}
 Suppose that we are given two subsets
  \[
              A
    \subseteq \{ \text{subsets $X \subseteq V$} \}
    \qquad\text{and}\qquad
              B
    \subseteq \{ \text{ideals $I \idealleq \mc{P}(V)$} \}
  \]
  such that the maps $\mc{I}, \mc{V}$ restrict to bijections $A \to B$ and $B \to A$.
  Then $A$ needs to be contained in the image of $\mc{V}$ while $B$ needs to be contained in the image of $\mc{I}$.
  It then follows that the bijections $A \to B$ and $B \to A$ are just restrictions of the mutually bijections from Corollary~\ref{corollary: bijection induced by Galois correspondence}.
  
  The $1$:$1$-correspondence given by Corollary~\ref{corollary: bijection induced by Galois correspondence} is therefore the most general $1$:$1$-correspondence which can be constructed between subsets of $V$ and ideal in $\mc{P}(V)$ by using the maps $\mc{V}, \mc{I}$.
  All other such correspondences must be restrictions of the one given by Corollary~\ref{corollary: bijection induced by Galois correspondence}.
  
  To better Corollary~\ref{corollary: bijection induced by Galois correspondence} we want to determine which kind of ideals can occur as vanishing ideals.
  We will show in the rest of this section that the vanishing ideals are precisely the radical ideals when $k$ is algebraically closed.
\end{fluff}


\begin{definition}
  Let $R$ be a commutative ring.
  \begin{enumerate}
    \item
      An ideal $I \idealleq R$ is a \emph{radical ideal} if for all $x \in R$, $n \geq 0$ it follows from $x^n \in I$ that $x \in I$.
    \item
      The \emph{radical} of an ideal $I \idealleq R$ is
      \[
                  \rad{I}
        \defined  \{
                    x \in R
                  \suchthat
                    \text{$x^n \in I$ for some $n \geq 0$}
                  \} \,.
      \]
  \end{enumerate}
\end{definition}


\begin{lemma}
  Let $R$ be a commutative ring and let $I \idealleq R$ be an ideal.
  Then $I$ is radical if and only if the ring $R/I$ is reduced, i.e.\ has no non-zero commutative elements.
\end{lemma}


\begin{proof}
  We have that
  \begin{align*}
        &\, \text{$R/I$ is reduced} \\
    \iff&\, \forall x \in R/I : (\text{$x^n = 0$ for some $n \geq 0$} \implies x = 0) \\
    \iff&\, \forall x \in R   : (\text{$\overline{x}^n = 0$ for some $n \geq 0$} \implies \overline{x} = 0) \\
    \iff&\, \forall x \in R   : (\text{$x^n \in I$ for some $n \geq 0$} \implies x \in I) \\
    \iff&\, \text{the ideal $I$ is radical} \,.
  \end{align*}
  This proves the claim.
\end{proof}


\begin{example}
  Let $R$ be a commutative ring.
  \begin{enumerate}
    \item
      The unit ideal $(1) = R$ is always radical.
    \item
      The zero ideal $0$ is radical if and only if $R = R/0$ is reduced.
    \item
      Every prime ideal $\mf{p} \ideallneq R$ is a radical ideal:
      
      If $x^n \in \mf{p}$ for some $x \in R$, $n \geq 0$ then $n \geq 1$ because $x^0 = 1 \notin \mf{p}$, and it then follows from $x^n \in \mf{p}$ and $\mf{p}$ being prime that $x \in \mf{p}$.
      
      Alternatively we can observe that $R/\mf{p}$ is an integral domain and is therefore reduced.
    \item
      It can conversely be shown that every radical ideal is an intersection of prime ideals.
  \end{enumerate}
\end{example}


% TODO (optional): More examples.


\begin{lemma}
  Let $R$ be a commutative ring and let $I \idealleq R$ be an ideal. 
  \begin{enumerate}
    \item
      \label{enumerate: radicals are radical}
      The radical $\rad{I}$ is a radical ideal in $R$.
    \item
      The radical $\rad{I}$ is the smallest radical ideal which contains $I$.
    \item
      For an ideal $J \idealleq R$ the following conditions are equivalent:
      \begin{enumerate}
        \item
          \label{enumerate: J is radical}
          The ideal $J$ is radical.
        \item
          \label{enumerate: J is some radical}
          There exists some ideal $I \idealleq R$ with $J = \rad{I}$.
        \item
          \label{enumerate: J is its own radical}
          The ideal $J$ satisfies $J = \rad{J}$.
      \end{enumerate}
  \end{enumerate}
\end{lemma}
\begin{proof}
    \leavevmode
  \begin{enumerate}
    \item
      We have that $0 \in \rad{I}$ because $0^1 = 0 \in I$.
      
      For $f, g \in \rad{I}$ there exist $n, m \geq 0$ with $f^n, g^m \in I$ and thus $f^s, g^t \in I$ for all $s \geq n, t \geq m$.
      It follows that
      \[
            (f + g)^{n+m}
        =   \sum_{\ell=0}^{n+m} \binom{n+m}{\ell} f^\ell g^{n+m-\ell}
        \in I
      \]
      because for all $i = 0, \dotsc, n+m$ we have that $\ell \geq n$ or $n+m-\ell \geq m$.
      This shows that also $f + g \in \rad{I}$.
      
      For $f \in \rad{I}$ there exists some $n \geq 0$ such that $f^n \in I$.
      For every $r \in R$ we then have that
      \[
            (rf)^n
        =   r^n f^n
        \in I
      \]
      and thus $rf \in \rad I$.
      
      For $x \in R$ with $x^n \in \rad{I}$ for some $n \geq 0$ there exists some $m \geq 0$ with $r^{mn} = (r^n)^m \in I$.
      It then follows that $x \in \rad{I}$.
      This shows that the ideal $\rad{I}$ is radical.
    \item
      We have that $I \subseteq \rad{I}$ because $x = x^1 \in I$ for every $x \in I$, and every radical ideal $J \idealleq R$ which contains $I$ must also contain the elements of $\rad{I}$.
    \item
      \begin{description}
        \item[\ref*{enumerate: J is radical} $\implies$ \ref*{enumerate: J is its own radical}:]
          The smallest radical ideal containing $J$ is just $J$ itself, so $J = \rad{J}$.
        \item[\ref*{enumerate: J is its own radical} $\implies$ \ref*{enumerate: J is some radical}:]
          Choose $I = J$.
        \item[\ref*{enumerate: J is some radical} $\implies$ \ref*{enumerate: J is radical}:]
          This follows from part~\ref*{enumerate: radicals are radical}.
        \qedhere
      \end{description}
  \end{enumerate}
\end{proof}


\begin{lemma}
  \label{lemma: vanishing ideals are radical}
  For every subset $X \subseteq V$ the ideal $\mc{I}(X)$ is a radical ideal in $\mc{P}(V)$.
\end{lemma}


\begin{proof}
  For $f \in \mc{P}(V)$ with $f^n \in \mc{I}(X)$ for some $n \geq 0$ we have that $f(x)^n = 0$ for every $x \in X$.
  Then $f(x) = 0$ for every $x \in X$ and thus $f \in \mc{I}(X)$.
\end{proof}


\begin{fluff}
  We will now show that the converse of Lemma~\ref{lemma: vanishing ideals are radical} holds if $k$ is algebraically closed.
  This will then answer our question which ideals $I \idealleq \mc{P}(V)$ are vanishing ideals:
  It is precisely the radical ideals.
  
  We proceed in three steps, each of which resulting in some kind of Nullstellensatz:
  We start off the weak Nullstellensatz, from which we then conclude the Nullstellensatz.
  By using the Rabinowitsch trick we then show the strong Nullstellensatz.
  
  To show the weak Nullstellensatz we will need a results from commutative algebra, namely Zariski’s lemma (Corollary~\ref{corollary: zariskis lemma}).
\end{fluff}


\begin{theorem}[Weak Nullstellensatz]
  \label{theorem: weak nullstellensatz}
  Let $k$ be algebraically closed.
  Then every maximal ideal $\mf{m} \subseteq k[X_1, \dotsc, X_n]$ is of the form
  \[
      \mf{m}
    = ( (X_1 - a_1), \dotsc, (X_n - a_n) )
    = \mf{m}_a
  \]
  for some $a = (a_1, \dotsc, a_n) \in k^n$.
\end{theorem}


\begin{proof}
  Let $R \defined k[X_1, \dotsc, X_n]$.
  The quotient $L \defined R/\mf{m}$ is a field because the ideal $\mf{m}$ is maximal and $L$ is finitely generated as a $k$-algebra because $R$ is a finitely generated $k$-algebra.
  It follows from \hyperref[corollary: zariskis lemma]{Zariski’s lemma} that the field extension $L/k$ is finite
  It follows that $L = k$ because $k$ is algebraically closed.
  
  Let $a_i \defined \overline{X_i} \in L$ for every $i = 1, \dotsc, n$.
  Then the ideal $\mf{m}_a = (X_1 - a_1, \dotsc, X_n - a_n)$ is a maximal (by Lemma~\ref{lemma: maximal ideal correspondin to a point}) with $\mf{m}_a \subseteq \mf{m}$.
  It follows that $\mf{m} = \mf{m}_a$.
\end{proof}


\begin{theorem}(Nullstellensatz)
  \label{theorem: nullstellensatz}
  Let $k$ be an algebraically closed field.
  For every proper ideal $I \ideallneq k[X_1, \dotsc, X_n]$ we have that $\mc{V}(I) \neq \emptyset$.
\end{theorem}


\begin{proof}
  There exists a maximal ideal $\mf{m} \ideallneq k[X_1, \dotsc, X_n]$ with $I \subseteq \mf{m}$ because $I$ is a proper ideal.
  By the weak Nullstellensatz we have that $\mf{m} = \mf{m}_a$ for some $a \in k^n$.
  We then have that $\{a\} = \mc{V}(\mf{m}_a) \subseteq \mc{V}(I)$.
\end{proof}


\begin{remark}
  \label{remark: (weak) nullstellensatz}
  \leavevmode
  \begin{enumerate}
    \item
      \label{enumerate: WNS follows from NS}
      The weak Nullstellensatz follows from the Nullstellensatz:
      If $\mf{m} \ideallneq k[X_1, \dotsc, X_n]$ is a maximal ideal then $\mf{m}$ is in particular a proper ideal, so by the Nullstellensatz there exists some $a \in k^n$ with $a \in \mc{V}(\mf{m})$.
      It follows that
      \[
                  \mf{m}
        \subseteq \mc{I}(\mc{V}(\mf{m}))
        \subseteq \mc{I}(a)
        =         \mf{m}_a
      \]
      and thus $\mf{m} = \mf{m}_a$ because both $\mf{m}, \mf{m}_a$ are maximal.
      
      The weak Nullstellensatz and Nullstellensatz are therefore equivalent.
    \item
      For the case $n = 1$ both the weak Nullstellensatz and the Nullstellensatz become a well-known characterization of algebraically closed fields:
      \begin{itemize}
        \item
          The maximal ideals of $k[X]$ are precisely the ideals of the form $(f)$ with $f \in k[X]$ irreducible because $k[X]$ is a principal ideal domain;
          the irreducible polynomial $f$ is unique if we require it to be monic.
          The weak Nullstellensatz therefore states for $n = 1$ that the irreducible monic polynomials in $k[X]$ are precisely the polynomials $X - a$ with $a \in k$.
        \item
          The proper ideal in $k[X]$ are precisely the ideals of the form $(f)$ with $f \in k[X]$ with $f = 0$ or $f$ being non-constant.
          The Nullstellensatz therefore states for $n = 1$ that the only polynomials $f \in k[x]$ with no roots are the non-zero constant ones.
      \end{itemize}
      This also shows that both Nullstellensätze hold only if $k$ is algebraically closed.
      
      Consider for example the case $k = \Real$ and $n = 1$.
      Then the ideal $(X^2 + 1) \idealleq \Real[X]$ is maximal but not of the form $X - a$ for some $a \in \Real$, and $\mc{V}((X^2 + 1)) = \emptyset$.
    \item
      \label{enumerate: partition of unity formulation of NS}
      The Nullstellensatz states that for all polynomials $f_1, \dotsc, f_s \in k[X_1, \dotsc, X_n]$ precisely one the following two things occurs:
      \begin{itemize}
        \item
          The polynomials $f_1, \dotsc, f_s$ have a common zero in $k^n$.
        \item
          There exists $g_1, \dotsc, g_n \in k[X_1, \dotsc, X_n]$ with $1 = g_1 f_1 + \dotsb + g_n f_n$.
      \end{itemize}
  \end{enumerate}
\end{remark}


\begin{theorem}[Strong Nullstellensatz]
  \label{theorem: strong nullstellensatz}
  If $k$ is algebraically closed then
  \[
      \mc{I}(\mc{V}(I))
    = \rad{I}
  \]
  for every ideal $I \idealleq k[X_1, \dotsc, X_n]$.
\end{theorem}
\begin{proof}
  The ideal $\mc{I}(\mc{V}(I))$ is radical by Lemma~\ref{lemma: vanishing ideals are radical} and contains $I$ by Lemma~\ref{lemma: properties of V and I} so it follows that $\rad{I} \subseteq \mc{I}(\mc{V}(I))$.
  
  To show the other inclusion let $h \in \mc{I}(\mc{V}(I))$.
  We need to show that $h^m \in I$ for some $m \geq 0$.
  For $h = 0$ this is clear, so we assume that $h \neq 0$.
  Let $f_1, \dotsc, f_s \in I$ with $I = (f_1, \dotsc, f_s)$;
  such $f_i$ exist because  $k[X_1, \dotsc, X_n]$ is noetherian.
  
  We use the \emph{Rabinowitsch trick}:
  We adjoin a new variable $Y$ to $k[X_1, \dotsc, X_n]$ and get $k[X_1, \dotsc, X_n, Y]$.
  Then $k[X_1, \dotsc, X_n]$ as a subring of $k[X_1, \dotsc, X_n, Y]$ so we can evaluate the polynomials $f \in k[X_1, \dotsc, X_n]$ at points $x \in k^{n+1}$.
  If $f_i(x) = 0$ for every $1 \leq i \leq s$ then $h(x) = 0$ which shows that the polynomials $f_1, \dotsc, f_s, 1 - h Y$ have no common zeros.
  
  It follows from the Nullstellensatz (see Remark~\ref{remark: (weak) nullstellensatz}) that there exist coefficients $p_1 \ldots, p_s, q \in k[X_1, \dotsc, X_n, Y]$ with
  \[
      1
    = p_0 (1 - Y h) + p_1 f_1 + \dotsb + p_s f_s \,.
  \]
  We can consider this as an identity in $k(X_1, \dotsc, X_n, Y)$ and replace $Y$ by $1/h$, resulting in the equality
  \[
      1
    =   p_1 \left(X_1, \dotsc, X_n, \frac{1}{h} \right) f_1
      + \dotsb
      + p_s \left(X_1, \dotsc, X_n, \frac{1}{h} \right) f_s \,.
  \]
  By multiplying both sides of this equation by a high enough power of $h$ we get find that
  \[
        h^m
    =   q_1 f_1 + \dotsb + q_s f_s
    \in I
  \]
  for some $m \geq 0$ and polynomials $q_1, \dotsc, q_s \in k[X_1, \dotsc, X_n]$.
\end{proof}


\begin{remark}
  The Rabinowitsch trick can be understood an an applicaiton of localization, as explained in \cite{MO90666}:
  
  We want to show that some power of $h$ is contained in $I$, which amounts to showing that the residue class $\overline{h}$ is nilpotent in $k[X_1, \dotsc, X_n]/I$.
  This happens only if $(k[X_1, \dotsc, X_n]/I)_{\overline{h}} = 0$.
  We have that
  \begin{align*}
            (k[X_1, \dotsc, X_n]/I)_{\overline{h}}
    &\cong  (k[X_1, \dotsc, X_n]/I)[Y]/(\overline{h}Y - 1)  \\
    &\cong  k[X_1, \dotsc, X_n,Y]/(I, hY - 1) \,,
  \end{align*}
  so we need to show that $(I, hY - 1)$ is not a proper ideal of $k[X_1, \dotsc, X_n, Y]$.
  By the Nullstellensatz this hols true if the polynomials of this ideal have no common roots, which holds because $h \in I$.
\end{remark}


\begin{remark}
  The strong Nullstellensatz implies the Nullstellensatz:
  
  If $I \ideallneq k[X_1, \dotsc, X_n]$ is a proper ideal with $\mc{V}(I) = \emptyset$ then $\rad{I} = \mc{I}(\mc{V}(I)) = \mc{I}(\emptyset) = (1)$ and thus $1 \in \rad{I}$.
  But then $1^m \in I$ for some $m \geq 0$ and thus $1 \in I$, which contradicts $I$ being a proper ideal.
  
  Together with part~\ref*{enumerate: WNS follows from NS} of Remark~\ref{remark: (weak) nullstellensatz} this shows that all three forms of the Nullstellensatz are equivalent.
  There are also other equivalent theorems which are commonly known as \enquote{the Nullstellensatz}, but we will not encounter them here.
\end{remark}


% TODO (optional): Give a more geometric interpretation of the proof once I know how.


\begin{corollary}
  \label{corollary: vanishing ideals are precisely radical ideals}
  If $k$ is algebraically closed then an ideal $I \idealleq k[X_1, \dotsc, X_n]$ is a vanishing ideal if and only if it is a radical ideal.
\end{corollary}


\begin{proof}
  Every vanishing ideal is a radical ideal by Lemma~\ref{lemma: vanishing ideals are radical}, and every radical ideal is a vanishing ideal by the \hyperref[theorem: strong nullstellensatz]{strong Nullstellensatz}.
\end{proof}


\begin{corollary}
  \label{corollary: algebraically closed correspondence for algebraic subsets and points}
  Let $k$ be algebraically closed.
  Then the maps $\mc{I}, \mc{V}$ restrict to the following mutually inverse bijections:
  \[
    \begin{matrix}
        \left\{
          \begin{tabular}{c}
              algebraic subsets \\
              $X \subseteq V$
          \end{tabular}
        \right\}
      & \begin{tikzcd}[column sep = large]
            {}
            \arrow[shift left]{r}{\mc{I}}
          & {}
            \arrow[shift left]{l}{\mc{V}}
        \end{tikzcd}
      & \left\{
          \begin{tabular}{c}
            radical ideals \\
            $I \idealleq \mc{P}(V)$
          \end{tabular}
        \right\}
      \\
        {}
      & {}
      & {}
      \\
        \rotatebox[origin=c]{90}{$\subseteq$}
      & {}
      & \rotatebox[origin=c]{90}{$\subseteq$}
      \\
        {}
      & {}
      & {}
      \\
        \left\{
          \begin{tabular}{c}
            points $a \in V$
          \end{tabular}
        \right\}
      & \begin{tikzcd}[column sep = large]
            {}
            \arrow[shift left]{r}{\mc{I}}
          & {}
            \arrow[shift left]{l}{\mc{V}}
        \end{tikzcd}
      & \left\{
          \begin{tabular}{c}
            maximal ideals \\
            $\mf{m} \idealleq \mc{P}(V)$
          \end{tabular}
        \right\}
    \end{matrix}
  \]
\end{corollary}


\begin{proof}
  In the diagram of Corollary~\ref{corollary: general correspondence for algebraic subsets and points} we can replace \enquote{vanishing ideals} by \enquote{radical ideals} by Corollary~\ref{corollary: vanishing ideals are precisely radical ideals}, and we can replace \enquote{vanishing ideals [\dots] which are maximal} first by \enquote{radical ideals [\dots] which are maximal} and then by \enquote{maximal ideal} because every maximal ideal is already a radical ideal (because it is prime).
\end{proof}


\begin{remark}
  It follows from part~\ref*{enumerate: VIV = V} of Lemma~\ref{lemma: properties of V and I} that the composition
  \[
            \mc{I} \circ \mc{V}
    \colon  \{ \text{ideals $I \idealleq \mc{P}(V)$} \}
    \to     \{ \text{ideals $I \idealleq \mc{P}(V)$} \}
  \]
  is idempotent and monotone with $I \subseteq (\mc{I} \circ \mc{V})(I)$ for all $I \idealleq \mc{P}(V)$.
  The composition $\mc{I} \circ \mc{V}$ is therefore a closure operator.
  If $k$ is algebraically closed then we have shown in this section that $\mc{I} \circ \mc{V}$ is the radical-operator $\rad{\phantom{I}}$.
\end{remark}


% \begin{remark}
%   One can further enhance the above correspondences:
%   
%   An algebraic subset $X \subseteq V$ is \emph{irreducible} if it cannot be decomposed as $X = Y_1 \cup Y_2$ for two proper algebraic sets $Y_1, Y_2 \subseteq V$.
%   
%   One can show that $X$ is irreducible if and only if the correspondig vanishing ideal $\mc{I}(X)$ is prime.
%   With this, one gets the following correspondences:
%   \[
%     \begin{tikzcd}[column sep = huge]
%           \left\{
%             \begin{tabular}{c}
%                 algebraic subsets \\
%                 $X \subseteq V$
%             \end{tabular}
%           \right\}
%           \arrow[shift left, shorten <= 1.5pt, shorten >= 4pt]{r}[above, xshift = -1.25 pt]{\mc{I}}
%         & \left\{
%             \begin{tabular}{c}
%               radical ideals \\
%               $I \idealleq \mc{P}(V)$
%             \end{tabular}
%           \right\}
%           \arrow[shift left, shorten >= 1.5pt, shorten <= 4pt]{l}[below, xshift = -1.25 pt]{\mc{V}}
%         \\
%           \left\{
%             \begin{tabular}{c}
%                 irreducible \\
%                 algebraic subsets \\
%                 $X \subseteq V$
%             \end{tabular}
%           \right\}
%           \arrow[phantom]{u}[rotate=90]{\subseteq}
%           \arrow[shift left, shorten >= 6.2pt]{r}[above, xshift = -3.1pt]{\mc{I}}
%         & \left\{
%             \begin{tabular}{c}
%               prime ideals \\
%               $\mf{p} \ideallneq \mc{P}(V)$
%             \end{tabular}
%           \right\}
%           \arrow[phantom]{u}[rotate=90]{\subseteq}
%           \arrow[shift left, shorten <= 6.2pt]{l}[below, xshift = -3.1pt]{\mc{V}}
%         \\
%           \left\{
%             \begin{tabular}{c}
%               points $a \in V$
%             \end{tabular}
%           \right\}
%           \arrow[phantom]{u}[rotate=90]{\subseteq}
%           \arrow[shift left, shorten <= 12pt]{r}[above, xshift = 6pt]{\mc{I}}
%         & \left\{
%             \begin{tabular}{c}
%               maximal ideals \\
%               $\mf{m} \ideallneq \mc{P}(V)$
%             \end{tabular}
%           \right\}
%           \arrow[phantom]{u}[rotate=90]{\subseteq}
%           \arrow[shift left,shorten >= 12pt]{l}[below, xshift = 6pt]{\mc{V}}
%     \end{tikzcd}
%   \]
% \end{remark}




