\subsection{Simple Rings \& Weddeburn}


\begin{definition}
  The ring $R$ is simple if it is nonzero and $0, R$ are the only two-sided ideals of $R$, i.e.\ if $R$ contains precisely two two-sided ideals.
\end{definition}


\begin{example}
  \label{example: matrix algebra over skew field is simple}
  If $D$ is a division ring and $n \geq 1$ then $M_n(D)$ is simple.
  This follows from the following lemma:
\end{example}


\begin{lemma}
  For every $n \geq 1$ the map
  \begin{align*}
              \{ \text{two-sided ideals $I \idealleq R$} \}
    &\longto  \{ \text{two-sided ideals $J \idealleq \Mat_n(R)$} \} \,,
    \\
                  I
    &\longmapsto  \Mat_n(I)
  \end{align*}
  is a well-defined bijection.
\end{lemma}


\begin{proof}
  If $I \idealleq R$ is a two-sided ideal then the canonical projection $\pi \colon R \to R/I$ is a ring homomorphism.
  The induced ring homomorphism $\Mat_n(R) \to \Mat_n(R/I)$ has $\Mat_n(I)$ as its kernel, which is therefore a two-sided ideal in $\Mat_n(R)$.
  
  Let on the other hand $J \idealleq \Mat_n(R)$ be a two-sided ideal.
  For all $i, j = 1, \dotsc, n$ let
  \[
      I_{ij}
    = \{
        r \in R
      \suchthat
        \text{there exists a matrix $A \in J$ whose $ij$-th coefficient is $r$}
      \} \,.
  \]
  Then $I_{ij}$ is a two-sided ideal in $R$:
  The projection $\pi_{ij} \colon \Mat_n(R) \to R$ onto the $ij$-th coefficient is a homomorphism of both left and right $R$-modules and the two-sided ideal $J$ is both a left and right $R$-submodule of $\Mat_n(R)$.
  It follows that $\pi_{ij}(J)$ is both a left and right $R$-submodule of $R$, i.e.\ a two-sided ideal.
  
  By multiplying a matrix $A \in J$ with permutation matrices from the left and from the right we can move every coefficient of $A$ to every other position without leaving $J$.
  It follows that the ideal $I \defined I_{ij}$ does not depend on the position $i,j$, and we have have that $J = \Mat_n(I)$ by construction of $I$.
\end{proof}


\begin{warning}
  A simple ring $R$ is not necessarily simple as an $R$-module:
  The ring $\Mat_n(D)$ for a skew field $D$ and $n \geq 2$ is a counterexample.
\end{warning}


% \begin{proposition}
%   \label{proposition: when semisimple is simple}
%   If $R$ is semisimple then the following are equivalent:
%   \begin{enumerate}
%     \item
%       \label{enumerate: is simple}
%       The ring $R$ is simple.
%     \item
%       \label{enumerate: only one simple}
%       The ring $R$ has only one simple $R$-module up to isomorphism
%     \item
%       \label{enumerate: is a matrix ring}
%       We have that $R \cong \Mat_n(D)$ for some $n \in \Natural$ and skew field $D$.
%   \end{enumerate}
% \end{proposition}
% 
% 
% \begin{proof}
%   It follows from the \hyperref[theorem: artin wedderburn theorem]{theorem of Artin--Wedderburn} that
%   \[
%     R \cong \Mat_{n_1}(D_1) \times \dotsb \times \Mat_{n_r}(D_r)
%   \]
%   for $r = |\Irr(R)|$, $n_1, \dotsc, n_r \geq 1$ and skew fields $D_1, \dotsc, D_r$.
%   
%   \begin{description}
%     \item[\ref*{enumerate: is a matrix ring} $\implies$ \ref*{enumerate: is simple}]
%       This follows from Example~\ref{example: simple ring}.
%     \item[\ref*{enumerate: is simple} $\implies$ \ref*{enumerate: only one simple}]
%       The $\Mat_{n_i}(D_i)$ correspond to two-sided ideal in $R$.
%       For $r \geq 2$ it would follow that $R$ contains nonzero proper two-sided ideals.
%     \item[\ref*{enumerate: only one simple} $\implies$ \ref*{enumerate: is a matrix ring}]
%       We have that $r = 1$ and therefore that $R \cong \Mat_{n_1}(D_1)$.
%     \qedhere
%   \end{description}
% \end{proof}
% 
% 
% \begin{theorem}[Wedderburn]
%   \label{theorem: wedderburns theorem}
%   If $R$ is simple then the following are equivalent:
%   \begin{enumerate}
%     \item
%       \label{enumerate: is semisimple}
%       The ring $R$ is semisimple.
%     \item 
%       \label{enumerate: is left artian}
%       The ring $R$ is (left) artian.
%     \item
%       \label{enumerate: has minimal left ideal}
%       The ring $R$ has a minimal nonzero left ideal $I$.
%     \item
%       \label{enumerate: is matrix ring over skew field}
%       We have that $R \cong \Mat_n(D)$ for some $n \in \Natural$ and skew field $D$.
%   \end{enumerate}
%   The skew field $D$ is then unique up to isomorphism.
% \end{theorem}
% 
% 
% \begin{proof}
%   The uniqueness of $D$ follows from the \hyperref[theorem: artin wedderburn theorem]{theorem of Artin--Wedderburn}
%   \begin{description}
%     \item[\ref*{enumerate: is semisimple} $\iff$ \ref*{enumerate: is matrix ring over skew field}]
%       This is part of Proposition~\ref{proposition: when semisimple is simple}.
%     \item[\ref*{enumerate: is semisimple} $\implies$ \ref*{enumerate: is left artian}]
%       This is part of Corollary~\ref{corollary: semisimple rings are notherian artinian}.
%     \item[\ref*{enumerate: is left artian} $\implies$ \ref*{enumerate: has minimal left ideal}]
%       Starting with any nonzero left ideal $I_1 \idealleq R$ there would otherwise exist for every $n \geq 1$ an ideal $I_{n+1} \idealleq R$ with $I_{n+1} \subsetneq I_n$, resulting in a descreasing sequence of ideals which does not stablize.
%     \item[\ref*{enumerate: has minimal left ideal} $\implies$ \ref*{enumerate: is semisimple}]
%       The isotypical component $R_I$ is a two-sided ideal by Lemma~\ref{lemma: isotypical components are two sided ideals} and nonzero by assumption.
%       It follows that $R = R_I$ is semisimple.
%     \qedhere
%   \end{description}
% \end{proof}


\begin{warning}
  Not every simple ring is semisimple, despite its name:
\end{warning}


\begin{example}
  \label{example: simple but not semisimple}
  Let $\ringchar(k) = 0$.
  The (first) Weyl algebra $\weyl = \weyl_1$ from subsection~\ref{subsection: first weyl algebra} is simple but not semisimple:
  
  To show that $\weyl$ is not semisimple it sufficies by Corollary~\ref{corollary: semisimple rings are notherian artinian} to observe that $\weyl$ is not artinian:
  We have seen that $\weyl$ has a basis given by $X^n \del^m$ with $n, m \geq 0$.
  It then follows for every $i \geq 0$ that
  \[
      \weyl \del^i
    = \gen{ X^n \del^{m+i} \suchthat n, m \geq 0 }_k
    = \gen{ X^n \del^m \suchthat n \geq 0, m \geq i }_k  \,,
  \]
  which shows that
  \[
                \weyl
    =           \weyl \del^0
    \idealgneq  \weyl \del^1
    \idealgneq  \weyl \del^2
    \idealgneq  \dotsb
  \]
  is a strictly decreasing sequence of left ideals of $\weyl$.
  
  To show that $\weyl$ is simple let $I \idealleq \weyl$ be a two-sided ideal and let $f \in I$ be non-zero.
  We can then write $f$ uniquely as
  \[
      f
    = \sum_{m \geq 0} p_m \del^m
  \]
  with $p_m \in k[X] \subseteq \weyl$ for all $m \geq 0$.
  It follows from $\del X = X \del + 1$ by induction that 
  \[
      \del^n X
    = X \del^n + n \del^{n-1}
  \]
  for all $n \geq 0$.
  It follows that
  \begin{align*}
        f X - X f
    &=  \sum_{m \geq 0} \left( p_m \del^m X - X p_m \del^m \right)  \\
    &=  \sum_{m \geq 0} \left( p_m X \del^m + m p_m \del^{m-1} - p_m X \del^m \right)
     =  \sum_{m \geq 0} m p_m \del^{m-1} \,.
  \end{align*}
  For the maximal index $m \geq 0$ with $p_m \neq 0$ it follows that $m! p_m \in I$ and therefore that
  \[
    p_m \in I \,.
  \]
  This shows that $I$ already contains a non-zero polynomial from $k[X] \subseteq \weyl$.
  We have seen in subsection~\ref{subsection: first weyl algebra} (namely Equation~\eqref{equation: motivation skew polynomial ring}) that the formula $\del X = X \del + 1$ generalizes to
  \[
    \del p = p \del + p'
  \]
  for all $p \in k[X] \subseteq \weyl$, where $p'$ denotes the (formal) derivative of $p$.
  We thus have that
  \[
      \del p - p \del
    = p'
  \]
  for all $p \in k[X]$.
  It follows for $d \defined \deg_X(p_m)$ that $p^{(d)} \in I$ is a non-zero constant polynomial (because $\ringchar(k) \neq 0$).
  This shows that $I$ contains a unit, which shows that $I = \weyl$.
\end{example}


\begin{remark}
  We have given in Remark~\ref{remark: skew polynomial rings} a short overview how the Weyl algebra $\weyl$ can be understood as a skew polyonmial ring over the ground ring $k[X]$ with respect to the $k$-derivation $\del \colon k[X] \to k[X]$.
  
  It can be shown more generally that for a field of characteristic $\ringchar(k) = 0$ and a $k$-algebra $A$ a skew polynomial ring $A[y;\delta]$ (with respect to a $k$-derivation $\delta \colon A \to A$) is simple if and only if $A$ is \emph{$\delta$-simple} and $\delta$ is not an \emph{inner derivation};
  proofs of this can be found in \cite[Theorem~3.15]{Lam1991First} and \cite[Proposition~2.1]{NoncommutativeNoetherian} (where definitions of the above terms can also be found).
  If $\ringchar(k) = 0$ then it also holds for every simple $k$-algebra $A$ and every non-inner derivation $\delta \colon A \to A$ that the skew polynomial ring $A[y;\delta]$ is a nonartinian simple ring; a proof of this can be found in \cite[Corollary~3.16]{Lam1991First}.
  
  This gives a recipe for constructing simple rings which are not semisimple.
\end{remark}


\begin{example}
  \label{example: simple but not semisimple endorphism ring}
  The following example is taken from \cite[3.14]{Lam1991First}:
  Let $V$ be a countable infinite-dimensional $k$-vector space for some (skew) field $k$.
  Then
  \[
              I
    \defined  \{
                f \in \End(V)
              \suchthat
                \text{$f$ has finite rank}
              \}
  \]
  is a two-sided ideal because we have for all $g \in \End(V)$ and $f, f_1, f_2 \in I$ that
  \begin{gather*}
            \rank(f \circ g),
            \rank(g \circ f)
      \leq  \rank(f)
  \shortintertext{and}
            \rank(f_1 + f_2)
      \leq  \rank(f_1) + \rank(f_2) \,.
  \end{gather*}
  The two-sided ideal $I$ is already maxmimal:
  
  Suppose that $f \in \End(V)$ has infinite rank.
  Let $C_1$ be a direct complement of $\im(f)$ and let $C_2$ be a direct complement of $\ker(f)$.
  Then $\im(f)$ is countable infinite-dimensional, so there exists an endomorphism $g_1 \colon V \to V$ such that the restriction $\restrict{g_1}{\im(f)} \to V$ is an isomorphism.
  The endomorphism $f$ restricts to an isomorphism $C_2 \to \im(f)$, so $C_2$ is also countable infinite-dimensional.
  It follows that there exists an endomorphism $g_2 \colon V \to V$ which restricts to an isomorphism $V \to C_2$.
  The composition $g_1 \circ f \circ g_2 \colon V \to V$ is then an isomorphism.
  This shows that the two-sided ideal generated by $f$ is already $\End(V)$ itself.
  
  It follows from the maximality of $I$ that $\End(V)/I$ is simple.
  We now show that $\End(V)/I$ is not noetherian, from which it follows by Corollary~\ref{corollary: semisimple rings are notherian artinian} that $\End(V)/I$ is not semisimple.
  
  We choose a basis $(b_{i,j})_{i,j \geq 0}$ of $V$, and for every $n \geq 0$ we consider the left ideal of $\End(V)$ given by
  \[
              J_n
    \defined  \{
                f \in \End(V)
              \suchthat
                \text{$f(b_{ij}) = 0$ for all $i \geq n$, $j \geq 0$}
              \} \,.
  \]
  We then have that $J_n \ideallneq J_{n+1}$ for all $n \geq 0$. 
  
  \begin{claim}
    For all $n \geq 0$ we have that $I + J_n \ideallneq I + J_{n+1}$.
  \end{claim}
  
  \begin{proof}
    It sufficies to show that $J_{n+1} \idealnleq I + J_n$:
    Consider an element $f \in J_{n+1}$ with
    \[
      f(b_{n+1, j}) = b_{n+1,j}
    \]
    for all $j \geq 0$.
    If $f \in I + J_n$ then there would exist $g \in I$, $f' \in J_n$ with $f = g + f'$.
    It then follows that
    \[
        b_{n+1,j}
      = f(b_{n+1,j})
      = g(b_{n+1,j}) + f'(b_{n+1,j})
      = g(b_{n+1,j})
    \]
    for all $j \geq 0$, and therefore that
    \[
        g(b_{n+1,j})
      = -b_{n+1,j}
    \]
    for all $j \geq 0$.
    But this contradicts $g$ having finite rank.
  \end{proof}

  It follows that
  \[
                I
    =           I + J_0
    \ideallneq  I + J_1
    \ideallneq  I + J_2
    \ideallneq  \dotsb
  \]
  is a strictly increasing sequence of ideals in $\End(V)$ and it follows that
  \[
                0
    =           (I + J_0)/I
    \ideallneq  (I + J_1)/I
    \ideallneq  (I + J_2)/I
    \ideallneq  \dotsb
  \]
  is a strictly increasing sequence of ideals in $\End(V)/I$.
  This shows that $\End(V)/I$ is not noetherian.
\end{example}


\begin{lemma}
  \label{lemma: isotypical components are two sided ideals}
  If $E$ is a simple $R$-module then the isotypical component $R_E$ is a two-sided ideal of $R$.
\end{lemma}


\begin{proof}
  The isotypical component $R_E$ is a submodule of $R$ and therefore a left ideal.
  For every $r \in R$ the map $R \to R$, $x \mapsto xr$ is a homomorphism of $R$-modules and therefore maps $R_E$ into $R_E$ by Lemma~\ref{lemma: functoriality of isotypical components}.
  This shows that $R_E$ is also a right ideal.
\end{proof}


\begin{theorem}[Wedderburn]
  \label{theorem: wedderburns theorem}
  The following conditions are equivalent:
  \begin{enumerate}
    \item
      \label{enumerate: simple and artinian}
      The ring $R$ is simple and (left) artinian.
    \item
      \label{enumerate: simple and minimal left ideal}
      The ring $R$ is simple and contains a minimal nonzero left ideal $I \idealleq R$.
    \item
      \label{enumerate: simple and semisimple}
      The ring $R$ is both simple and semisimple.
    \item
      \label{enumerate: semisimple with unique simple}
      The ring $R$ is semisimple and has only one simple module up to isomorphism.
    \item
      \label{enumerate: matrix algebra over skew field}
      We have that $R \cong \Mat_n(D)$ for some $n \geq 1$ and skew field $D$.
  \end{enumerate}
  The skew field $D$ is then up to isomorphism uniquely determined as $D \cong \End_R(I)^\op$, with $I$ as above being the unique simple $R$-module up to isomorphism.
  The number $n$ is uniquely determined as the multiplicity of $I$ in $R$.
\end{theorem}


\begin{proof}
  \label{proof: wedderburns theorem first proof}
  \leavevmode
  \begin{description}
    \item[\ref*{enumerate: simple and artinian} $\implies$ \ref*{enumerate: simple and minimal left ideal}]
      The set of nonzero left ideals of $R$ is non-empty because $R \neq 0$ and thus contains a minimal element because $R$ is artinian.
    \item[\ref*{enumerate: simple and minimal left ideal} $\implies$ \ref*{enumerate: simple and semisimple}]
      The ideal $I$ is a simple $R$-module and the $I$-isotypical component $R_I$ is a nonzero two-sided ideal by Lemma~\ref{lemma: isotypical components are two sided ideals}.
      It follows from $R$ being simple that $R_I = R$, and therefore that $R$ is semisimple.
    \item[\ref*{enumerate: simple and semisimple} $\implies$ \ref*{enumerate: semisimple with unique simple}]
      Because $R \neq 0$ is a sum of simple $R$-modules it follows that there exists a simple $R$-module $E$.
      The $E$-isotypical component $R_E$ is a two-sided ideal by Lemma~\ref{lemma: isotypical components are two sided ideals} which is nonzero by Lemma~\ref{lemma: simple module of semisimple ring is direct summand}.
      It follows that $R_E = R$, and therefore from Lemma~\ref{lemma: simple module of semisimple ring is direct summand} that $E$ is the unique simple $R$-module up to isomorphism.
    \item[\ref*{enumerate: semisimple with unique simple} $\implies$ \ref*{enumerate: matrix algebra over skew field}]
      This follows from the \hyperref[theorem: artin wedderburn theorem]{theorem of Artin--Wedderburn}.
    \item[\ref*{enumerate: matrix algebra over skew field} $\implies$ \ref*{enumerate: simple and semisimple}]
      That $\Mat_n(D)$ is simple follows from Example~\ref*{example: matrix algebra over skew field is simple}, and that $R$ is semisimple follows from Example~\ref{example: semisimple rings}.
    \item[\ref*{enumerate: simple and semisimple} $\implies$ \ref*{enumerate: simple and artinian}]
      This follows from Corollary~\ref{corollary: semisimple rings are notherian artinian}.
  \end{description}
  The minimal ideal $I$ is a simple submodule of $R$, and thus the unique simple $R$-module up to isomorphism .
  Because $D^n$ is a simple $\Mat_n(D)$-module, so it follows that
  \[
          D^\op
    \cong \End_{\Mat_n(D)}(D^n)
    \cong \End_R(I) \,,
  \]
  and therefore that $D \cong \End_R(I)^\op$.
  The multiplicity of $I$ in $R$ is the same as the multiplicity of $D^n$ in $\Mat_n(D)$, which is $n$.
\end{proof}




