\chapter{Extension of scalars}
For this chapter we fix a field extension $L/k$.




\section{Definition and universal property}


\begin{defi}
 For a $k$-vector space $V$ we define the $L$-vector spaces $V_L$ as
 \[
  V_L := L \otimes_k V
 \]
 where the multiplication is defined as
 \[
  \lambda \cdot (l \otimes v) = (\lambda l) \otimes v
 \]
 for all $\lambda \in L$ and simple tensors $l \otimes v \in V_L$.
\end{defi}

To see that this multiplication is well-defined notice that for every $\lambda \in L$ the multiplication with $\lambda$ in $L$, namely
\[
 \pi_\lambda \colon L \to L, l \mapsto \lambda l,
\]
is $L$-linear and thus $k$-linear. The multiplication with $\lambda \in L$ as defined above is simply $\pi_\lambda \otimes \id_V$.


For any $k$-vector space $V$ we have a $k$-linear inclusion
\[
 \can_V \colon V \hookrightarrow V_L, v \mapsto 1 \otimes v.
\]
The $L$-vector space $V_L$ together with the inclusion $\can_V$ satisfies an universal property.


\begin{thrm}[Universal property of the extension of scalars]
 Let $V$ be a $k$-vector space. Then for every $L$-vector space $W$ we have a natural isomorphism of $k$-vector spaces
 \begin{align*}
             \beta_W \colon \Hom_L(V_L, W) &\to \Hom_k(V,W), \\
                                         f &\mapsto f \circ \can_V, \\
  (\lambda \otimes v \mapsto \lambda g(v)) &\mapsfrom g
 \end{align*}
 This property defines $V_L$ uniquely up to unique isomorphsm: If $V'$ is an $L$-vector space and $\iota \colon V \to V'$ a $k$-linear map such that
 \[
  \alpha_W \colon \Hom_L(V', W) \to \Hom_k(V,W), f \mapsto f \circ \iota
 \]
 is an isomorphism of $k$-vector spaces for every $L$-vector space $W$ then there exists a unique isomorphism of $L$-vector spaces
 \[
  \phi \colon V_L \to V'
 \]
 such that the diagram
 \begin{center}
  \tikzsetnextfilename{scalar_extension_universal_property}
  \begin{tikzpicture}[node distance = 6em, auto]
   \node (V) {$V$};
   \node (V_L) [below left of = V] {$V_L$};
   \node (V')  [below right of = V] {$V'$};
   \draw[->] (V) to node[swap] {$\can_V$} (V_L);
   \draw[->] (V) to node {$\iota$} (V');
   \draw[->] (V_L) to node {$\phi$} (V');
  \end{tikzpicture}
 \end{center}
 commutes.
\end{thrm}


That the $\beta_W$ are ``natural'' isomorphisms means the following: Fixing a $k$-vector space $V$ we have two functors
\[
 F_1, F_2 \colon \cVect{L} \to \cVect{k}
\]
where $F_1(W) := \Hom_L(V_L,W)$ for every $L$-vector space $W$ and for every $L$-linear map $f \colon W \to W'$
\[
 F_1(f) \colon \Hom_L(V_L, W) \to \Hom_L(V_L, W'), h \mapsto f \circ h,
\]
as well as $F_2(W) := \Hom_k(V,W)$ for every $L$-vector space $W$ and for every $L$-linear map $f \colon W \to W'$
\[
 F_2(f) \colon \Hom_k(V,W) \to \Hom_k(V,W'), h \mapsto f \circ h.
\]
The claim is that the $\beta_W$ form an natural isomorphism from $F_1$ to $F_2$, i.e.\ they form a natural transformation from $F_1$ to $F_2$ and $\beta_W$ is an isomorphism for every $L$-vector space $W$.


\begin{proof}
 We will start by proving that $\beta_W$ is an isomorphism of $k$-vector spaces for every $L$-vector space $W$. For this we fix an $L$-vector space $W$. It is clear that $\Phi := \beta_W$ is well-defined and $k$-linear. To show that $\Phi$ is an isomorphism we show that
 \[
  \Psi \colon \Hom_k(V,W) \to \Hom_L(V_L, W), g \mapsto \left( (\lambda \otimes v) \mapsto \lambda g(v) \right).
 \]
 is an inverse of $\Phi$. That $\Psi(g)$ is well-defined and $k$-linear for every map \mbox{$g \in \Hom_k(V,W)$} follows from the fact that the map
 \[
  L \times V \to W, (\lambda, v) \mapsto \lambda g(v)
 \]
 is well-defined and $k$-bilinear. That $\Psi(g)$ is also $L$-linear is clear. That $\Psi$ itself is also $k$-linear is also clear.
 
 Since for all $f \in \Hom_L(V_L,W), \lambda \in L, v \in V$
 \begin{align*}
   &\,(\Psi \Phi)(f)(\lambda \otimes v) \\
  =&\,\Psi(\Phi(f))(\lambda \otimes v)
  =   \Psi(f \circ \can_V)(\lambda \otimes v) \\
  =&\,\lambda (f \circ \can_V)(v)
  =   \lambda f(\can_V(v)) \\
  =&\,\lambda f(1 \otimes v)
  =   f (\lambda \otimes v)
 \end{align*}
 we have $\Psi \Phi = \id_{\Hom_L(V_L, W)}$ and since for all $g \in \Hom_k(V,W), v \in V$
 \begin{align*}
  (\Phi \Psi)(g)(v)
  &= \Phi(\Psi(g))(v)
  = (\Psi(g) \circ \can_V)(v) \\
  &= \Psi(g)(\can_V(v))
  = \Psi(g)(1 \otimes v) \\
  &= g(v)
 \end{align*}
 we have $\Phi \Psi = \id_{\Hom_k(V,W)}$.
 
 Next we will show that the $\beta_W$ are natural isomorphisms. For this we need to check that for all $L$-vector spaces $W$ and $W'$ and every $L$-linear map \mbox{$f \colon W \to W'$} the diagram
 \begin{center}
  \tikzsetnextfilename{naturalisomorphism}
  \begin{tikzpicture}[node distance = 6em, auto]
   \node (HomLW) {$\Hom_L(V_L, W)$};
   \node (HomLW') [right = 6em of HomLW] {$\Hom_L(V_L, W')$};
   \node (HomkW) [below of = HomLW] {$\Hom_k(V,W)$};
   \node (HomkW') [below of = HomLW'] {$\Hom_k(V,W')$};
   \draw[->] (HomLW) to node {$F_1(f)$} (HomLW');
   \draw[->] (HomkW) to node[swap] {$F_2(f)$} (HomkW');
   \draw[->] (HomLW) to node[swap] {$\beta_W$} (HomkW);
   \draw[->] (HomLW') to node {$\beta_{W'}$} (HomkW');
  \end{tikzpicture}
 \end{center}
 commutes, where $F_1$ and $F_2$ are defined as in the previous explanation. This is clear, since $F_1(f)$ and $F_2(f)$ are precomposing with $f$ and $\beta_W$ and $\beta_{W'}$ are composing with $\can_V$, and therefore for all $h \in \Hom_L(V_L, W)$
 \begin{align*}
   &\, (F_2(f) \circ \beta_W)(h)
  =    F_2(f)(\beta_W(h)) \\
  =&\, F_2(f)(h \circ \can_V)
  =    f \circ (h \circ \can_V) \\
  =&\, (f \circ h) \circ \can_V
  =    \beta_{W'}(f \circ h) \\
  =&\, \beta_{W'}(F_1(f)(h))
  =    (\beta_{W'} \circ F_1(f))(h).
 \end{align*}
 
 All that’s left to show is that property defines $V_L$ uniquely up to unique isomorphism. To prove this let $V'$ be an $L$-vector space together with a $k$-linear map $\iota \colon V \to V'$ such that for every $L$-vector space $W$ the map
 \[
  \alpha_W \colon \Hom_L(V', W) \to \Hom_k(V, W), f \mapsto f \circ \iota
 \]
 is an isomorphism of $k$-vector spaces.
 
 First we notice that
 \[
  \Hom_L(V_L, V_L) \cong \Hom_k(V, V_L), f \mapsto f \circ \can_V.
 \]
 Therefore there exists exactly one $f \in \Hom_L(V_L, V_L)$ with $f \circ \can_V = \can_V$. It is clear that $f = \id_{V_L}$. In the same way we find that $g \in \Hom_L(V', V')$ with $g \circ \iota = \iota$ if and only if $g = \id_{V'}$.
 
 Next we notice that
 \[
  \Hom_L(V_L, V') \cong \Hom_k(V, V'), f \mapsto f \circ \can_V.
 \]
 Therefore there exists a unique $L$-linear map $\phi \in \Hom_L(V_L, V')$ such that \mbox{$\iota = \phi \circ \can_V$}. All that’s left to check is that $\phi$ is an isomorphism. For this we find in same way that there exists a unique $L$-linear map $\psi \in \Hom_L(V', V_L)$ with $\can_V = \psi \circ \iota$. Since
 \[
  \phi \psi \circ \iota = \phi \circ \can_V = \iota
 \]
 as well as
 \[
  \psi \phi \circ \can_V= \psi \circ \iota = \can_V
 \]
 we find by the previous observations that $\phi \psi = \id_{V'}$ and $\psi \phi = \id_{V_L}$.
\end{proof}


Recall the following from linear algebra:


\begin{rec}
 Let $k$ be a field and $V$ a $k$-vector space. Then $B \subseteq V$ is a $k$-basis of $V$ if and only if for every $k$-vector space $W$ the restriction
 \[
  \Hom_k(V,W) \to \Maps(B,W), f \mapsto f_{|B}
 \]
 is a bijection.
\end{rec}


\begin{lem}
 Let $V$ be a $k$-vector space. Then $\{v_i\}_{i \in I}$ is a $k$-basis of $V$ if and only if $\{1 \otimes v_i\}_{i \in I}$ is an $L$-basis of $V_L$.
\end{lem}
\begin{proof}
 Since $\can_V \colon V \to V_L$ is injective it induces a bijection
 \[
  \phi \colon \{v_i\}_{i \in I} \to \{1 \otimes v_i\}_{i \in I}, v_i \mapsto 1 \otimes v_i = \can_V(v_i)
 \]
 For every $L$-vector space $W$ we therefore get a commutative diagram
 \begin{center}
  \tikzsetnextfilename{scalar_extension_and_bases}
  \begin{tikzpicture}[node distance = 6em, auto]
   \node (HomL) {$\Hom_L(V_L, W)$};
   \node (Homk) [right = 6em of HomL] {$\Hom_k(V,W)$};
   \node (MapsL) [below of = HomL] {$\Maps\left( \{1 \otimes v_i\}_{i \in I}, W\right )$};
   \node (Mapsk) [below of = Homk] {$\Maps\left( \{v_i\}_{i \in I} ,W\right )$};
   \draw[double equal sign distance] (HomL) to node {$\sim$} (Homk);
   \draw[double equal sign distance] (MapsL) to node {$\sim$} (Mapsk);
   \draw[->] (HomL) to node[swap] {restriction} (MapsL);
   \draw[->] (Homk) to node {restriction} (Mapsk);
  \end{tikzpicture}
 \end{center}
 where we have an isomorphism of $k$-vector spaces
 \[
  \Hom_L(V_L, W) \cong \Hom_k(V, W), h \mapsto h \circ \can_V,
 \]
 and the bijection
 \[
  \Maps\left( \{1 \otimes v_i\}_{i \in I}, W\right ) \cong \Maps\left( \{v_i\}_{i \in I} ,W\right ), h \mapsto h \circ \phi.
 \]
 That $\{v_i\}_{i \in I}$ is a $k$-basis of $V$ is equivalent to saying that the restriction on the right side is a bijection and that $\{1 \otimes v_i\}_{i \in I}$ is an $L$-basis of $V_L$ is equivalent to saying that the restriction on the left side is a bijection. Since the diagram commutes these two are equivalent.
\end{proof}


\begin{cor}\label{cor: inclusion to bijection vector spaces}
 Let $V$ be an $L$-vector space, $U \subseteq V$ a $k$-vector subspace and $B \subseteq U$, such that $B$ is a $k$-basis of $U$ and an $L$-basis of $V$. Then
 \[
  \phi \colon U_L \colon V, \lambda \otimes u \mapsto \lambda u
 \]
 is an isomorphism of $L$-vector spaces.
\end{cor}
\begin{proof}
 $\phi$ is the $L$-linear map that corresponds to the $k$-linear inclusion $U \hookrightarrow V$. Therefore it maps the $L$-basis $\{1 \otimes b\}_{b \in B}$ of $U_L$ bijectively to the $L$-basis $B$ of $V$.
\end{proof}





\section{Properties as a functor}


Given $k$-vector spaces $V$ and $W$ and a $k$-linear map $f \colon V \to W$ we get a $k$-linear map
\[
 f_L := \id_L \otimes f \colon V_L \to W_L.
\]
It is easy to check that $f_L$ is also $L$-linear. Therefore we can understand the extension of scalars as a functor
\[
 E \colon \cVect{k} \to \cVect{L}
\]
with $E(V) = V_L$ for every $k$-vector space $V$ and $E(f) = f_L$ for every $k$-linear map $f$. 

\begin{lem}
 Let $V$ and $V'$ be $k$-vector spaces. For every $k$-linear map $f : V \to V'$
 \tikzsetnextfilename{can_and_fL_commute}
 \begin{center}
  \begin{tikzpicture}[node distance = 6em, auto]
   \node (V) {$V$};
   \node (V') [right of = V] {$V'$};
   \node (VL) [below of = V] {$V_L$};
   \node (V'L) [below of = V'] {$V'_L$};
   \draw[->] (V) to node {$f$} (V');
   \draw[->] (VL) to node {$f_L$} (V'L);
   \draw[->] (V) to node[swap] {$\can_V$} (VL);
   \draw[->] (V') to node {$\can_{V'}$} (V'L);
  \end{tikzpicture}
 \end{center}
 commutes.
\end{lem}
\begin{proof}
 This can easily be checked by calculation, since for all $v \in V$
 \begin{align*}
  (f_L \circ \can_V)(v)
  &= f_L(\can_V(v))
  = f_L(1 \otimes v) \\
  &= 1 \otimes f(v)
  = \can_{V'}(f(v))
  = (\can_{V'} \circ f)(v).
 \end{align*}
\end{proof}


We also have the restriction of scalars as a functor
\[
 R \colon \cVect{L} \to \cVect{k}
\]
which sends every $L$-vector space to its underlying $k$-vector space and every $L$-linear map to the corresponding $k$-linear map.
 
From the universal property of the extension of scalars we know that for every $k$-vector space $V$ and $L$-vector space $W$ we have an isomorphism
\[
 \Phi_{V,W} \colon \Hom_L(E(V),W) \cong \Hom_k(V,R(W))
\]
of $k$-vector spaces. This observation results in the following proposition:
 
\begin{prop}
 $E$ is left adjoint to $R$ via the bijections $\Phi_{V,W}$.
\end{prop}
\begin{proof}
 We need to check that for all $k$-vector spaces $V$ and $V'$, all $L$-vector spaces $W$ and $W'$, every $k$-linear map $f \colon V' \to V$ (notice that $f$ goes from $V'$ to $V$) and every $L$-linear map $g \colon W \to W'$ the diagram
 \tikzsetnextfilename{scalar_extension_and_restriction_adjoint}
 \begin{center}
  \begin{tikzpicture}[node distance = 6em, auto]
   \node (L) {$\Hom_L(V_L, W)$};
   \node (L') [right = 6em of L] {$\Hom_L(V'_L, W')$};
   \node (k) [below of = L] {$\Hom_k(V,W)$};
   \node (k') [below of = L'] {$\Hom_k(V', W')$};
   \draw[->] (L) to node {$g \circ - \circ f_L$} (L');
   \draw[->] (k) to node[swap] {$g \circ - \circ f$} (k');
   \draw[->] (L) to node[swap] {$\Phi_{V,W}$} (k);
   \draw[->] (L') to node {$\Phi_{V',W'}$} (k');
  \end{tikzpicture}
 \end{center}
 commutes. Since $\Phi_{V,W} = - \circ \can_V$ and $\Phi_{V,W'} = - \circ \can_{V'}$ this follows from $\can_{V'} \circ f_L = f \circ \can_V$.
\end{proof}





\section{Extension of scalars for algebras}


Given $k$-algebras $A_1$ and $A_2$ their tensor product $A_1 \otimes_k A_2$ is a $k$-algebra via the multiplication
\[
 (a_1 \otimes b_1) \cdot (a_2 \otimes b_2) = (a_1 a_2) \otimes (b_1 b_2).
\]
To see that this multiplication is well-defined notice that the map
\begin{align*}
 A_1 \times A_2 \times A_1 \times A_2 &\to A_1 \otimes_k A_2, \\
 (a_1, b_1, a_2, b_2) &\mapsto (a_1 a_2) \otimes (b_1 b_2)
\end{align*}
is well-defined and $k$-multilinear, so we get a $k$-linear map
\begin{align*}
 A_1 \otimes_k A_2 \otimes_k A_1 \otimes_k A_2 &\to A_1 \otimes_k A_2, \\
 a_1 \otimes b_1 \otimes a_2 \otimes b_2 &\mapsto (a_1 a_2) \otimes (b_1 b_2),
\end{align*}
which under the isomorphism of $k$-vector spaces
\[
 A_1 \otimes_k A_2 \otimes_k A_1 \otimes_k A_2 \cong (A_1 \otimes_k A_2) \otimes_k (A_1 \otimes_k A_2)
\]
corresponds to a $k$-bilinear map
\begin{align*}
 (A_1 \otimes_k A_2) \times (A_1 \otimes_k A_2) &\to A_1 \otimes_k A_2, \\
 (a_1 \otimes b_1, a_2 \otimes b_2) &\mapsto (a_1 a_2) \otimes (b_1 b_2).
\end{align*}

Given a $k$-algebra $A$ we get that $A_L$ is an $L$-vector space and a $k$-algebra. It is easy to see that that these structures are compatible and give $A_L$ the structure of an $L$-algebra. To see this compatibility simply notice that the $k$-bilinear map
\begin{align*}
 (A_1 \otimes_k A_2) \times (A_1 \otimes_k A_2) &\to A_1 \otimes_k A_2, \\
 (a_1 \otimes b_1, a_2 \otimes b_2) &\mapsto (a_1 a_2) \otimes (b_1 b_2).
\end{align*}
is also $L$-bilinear.

\begin{lem}
 For $k$-algebras $A_1$ and $A_2$ and a homomorphism of $k$-algebras
 \[
  f \colon A_1 \to A_2
 \]
 the corresponding $L$-linear map
 \[
  f_L \colon L \otimes_k A_1 \to L \otimes_k A_2
 \]
 is a homomorphism of $L$-algebras.
\end{lem}
\begin{proof}
 $f_L$ is multiplicative on the simple tensors since for all simple tensors $\lambda_1 \otimes a_1, \lambda_2 \otimes a_2 \in L \otimes_k A_1$
 \begin{align*}
   &\, f_L((\lambda_1 \otimes a_1)(\lambda_2 \otimes a_2)) \\
  =&\, f_L((\lambda_1 \lambda_2) \otimes (a_1 a_2))
  =    (\lambda_1 \lambda_2) \otimes f(a_1 a_2) \\
  =&\, (\lambda_1 \lambda_2) \otimes (f(a_1)f(a_2))
  =    (\lambda_1 \otimes f(a_1)) (\lambda_2 \otimes f(a_2)) \\
  =&\, f_L(\lambda_1 \otimes a_1) f_L(\lambda_2 \otimes a_2).
 \end{align*}
 Since these simple tensors generate $L \otimes_k A_1$ it follows that $f_L$ is multiplicative.
\end{proof}
 
This shows that we can understand the extension of scalars for $k$-algebras als a functor from $\cAlg{k}$ to $\cAlg{L}$.

\begin{lem}
 Let $A$ be a $k$-algebra, $B$ an $L$-algebra and $f \colon A \to B$ a homomorphism of $k$-algebras. Then the corresponding $L$-linear map $\hat{f} \colon A_L \to B$ is a homomorphism of $L$-algebras.
 \begin{proof}
 Notice that for all simple tensors $\lambda_1 \otimes a_1, \lambda_2 \otimes a_2 \in A_L$ we have
  \begin{align*}
   \hat{f}((\lambda_1 \otimes a_1)(\lambda_2 \otimes a_2))
   &= \hat{f}((\lambda_1 \lambda_2) \otimes (a_1 a_2))
   = \lambda_1 \lambda_2 f(a_1 a_2) \\
   &= \lambda_1 \lambda_2 f(a_1) f(a_2)
   = \lambda_1 f(a_1) \lambda_2 f(a_2) \\
   &= \hat{f}(\lambda_1 \otimes a_1) \hat{f}(\lambda_2 \otimes a_2).
  \end{align*}
  Since the simple tensors generate $A_L$ it follows that $\hat{f}$ is multiplicative.
 \end{proof}
\end{lem}


\begin{cor}\label{cor: inclusion to bijection algebras}
 Let $B$ be an $L$-algebra, $A \subseteq B$ a $k$-subalgebra and $X \subseteq A$, such that $X$ is a $k$-basis of $A$ and an $L$-basis of $B$. Then
 \[
  \phi \colon A_L \to B, \lambda \otimes a \mapsto \lambda a
 \]
 is an isomorphism of $L$-algebras.
\end{cor}
\begin{proof}
 We know that $\phi$ is a homomorphism of $L$-algebras and we have already seen that $\phi$ is an isomorphism of $L$-vector spaces.
\end{proof}


\begin{lem}
 Let $A$ be a $k$-algebra and $I \subseteq A$ a left-ideal. Then $I_L$ is an left-ideal in the $L$-algebra $A_L$.
 %TODO: Furthermore, if $I$ is generated by $(b_j)_{j \in J}$ as an left ideal in $A$, then $I_L$ is generated by $(1 \otimes b_j)_{j \in J}$ as an left ideal in $A_L$.
\end{lem}
\begin{proof}
 Since $I$ is a $k$-vector subspace of $A$ we know that $I_L$ is an $L$-vector subspace of $A_L$. For $x \in A_L$ and $y \in I_L$ we have
 \[
  x = \sum_{i=1}^n \lambda_i \otimes a_i
 \]
 with $\lambda_i \in L$ and $a_i \in A$ and
 \[
  y = \sum_{j=1}^m \lambda'_j \otimes a'_j
 \]
 with $\lambda'_j \in L$ and $a'_j \in I$, and therefore
 \begin{align*}
  xy
  &= \left( \sum_{i=1}^n \lambda_i \otimes a_i \right) \left( \sum_{j=1}^m \lambda'_j \otimes a'_j \right) \\
  &= \sum_{i=1}^n \sum_{j=1}^m (\lambda_i \lambda'_j) \otimes (a_i a'_j)
  \in I_L
 \end{align*}
 This shows that $I_L$ is a left ideal in $A_L$.
\end{proof}





\section{Polynomial rings and matrix rings}


For the polynomial ring $k[X_1, \dotsc, X_n]$ we have an $L$-algebra
\[
 k[X_1, \dotsc, X_n]_L = L \otimes_k k[X_1, \dotsc, X_n].
\]
The interesting thing about this algebra is that is it actually $L[X_1, \dotsc, X_n]$.


\begin{prop}
 We have an isomorphism of $L$-algebras
 \[
  \Phi \colon k[X_1, \dotsc, X_n]_L \to L[X_1, \dotsc, X_n], \lambda \otimes f \mapsto \lambda f.
 \]
\end{prop}
\begin{proof}
 The monomial $X^\alpha$ with $\alpha \in \N^n$ are a $k$-basis of $k[X_1, \dotsc, X_n]$ and an $L$-basis of $L[X_1, \dotsc, X_n]$ so it follows from corollary \ref{cor: inclusion to bijection algebras}.
\end{proof}


Another interesting example of the extension of scalars is $L \otimes_k \Mat_n(k)$.


\begin{prop}
 We have an isomorphism of $L$-algebras
 \[
  \Phi \colon L \otimes_k \Mat_n(k) \to \Mat_n(L).
 \]
\end{prop}
\begin{proof}
 The matrices $(E_{ij})_{1 \leq i,j \leq n}$ are a $k$-basis of $k[X_1, \dotsc, X_n]$ and an $L$-basis of $L[X_1, \dotsc, X_n]$ so it follows from corollary \ref{cor: inclusion to bijection algebras}.
\end{proof}





















